---
title: "strainphlan_allcurate_CRC_IBD_T2D"
output: html_document
date: "2025-04-06"

---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/Xiaoqian/Desktop/pop_gen/UHGG_plus4/strainphlan/allcurated_with_IBD_T2D_meta4_beta1_vJan21')
```

##setup
```{r libararies, include=FALSE}
library(phangorn)
library(ggtree)
library(ggplot2)
library(castor)
library(RColorBrewer)
library(dplyr)
library(ggnewscale)
library(reshape2)
library(purrr)
library(ComplexHeatmap)
library(tidyr)
library(SignifReg)
library(tibble)
library(caret)
library(scales)
library(forcats)
library(phytools)
library(ggtext)
library(wesanderson)

```

```{r source, echo=FALSE}
source('../merge4d_meta4_beta1_vJan21/sweep_finder.R')
```

#import tree data and call sweeps
```{r import data, echo=FALSE}
folder_list_all <- Sys.glob("*_marker/")
tree_list_all=lapply(folder_list_all, function(x) Sys.glob(paste0(x,'RAxML_bestTree*.tre')))

taxa_list=read.table("../../fastANI_clust.average.diffH.94s.fgspecies.txt",
                     sep='\t',header=T)
taxa_list_summary=taxa_list %>% group_by(cluster) %>% count(fgspecies) %>% slice_max(order_by=n,with_ties = F)
taxa_list_summary$species=sapply(strsplit(taxa_list_summary$fgspecies,split = ';'),'[[',3)
taxa_list_summary$genus=sapply(strsplit(taxa_list_summary$fgspecies,split = ';'),'[[',2)
taxa_list_summary=taxa_list_summary %>% mutate(taxa = case_when(species=='s__' ~ paste0(genus,'_',species),
                                                                TRUE ~ paste0(species)))
taxa_list_summary$cluster_taxa=paste(taxa_list_summary$taxa,
                                     taxa_list_summary$cluster,sep = "_")
taxa_list_summary=taxa_list_summary %>% arrange(genus)

sweep_list_all=read.csv('../../instrain/confirmed_sweep_from_pairwise.csv')
df_strainphlan_cutoff=read.csv('../../SGBdiffH_500_1pois_1nb_finda/strainphlan_cutoff_new.csv')

sweep_list_all$SGB=sapply(strsplit(sweep_list_all$scaffold,split = 'diffH'),'[[',1)
sweep_list_all=left_join(sweep_list_all,taxa_list_summary,by=c("SGB"="cluster"))


```

```{r}
rf_list=list()
sf_list=list()
newtotalsweep_list=list()
newtotalstrain_list=list()
for (i in 1: length(folder_list_all)){
  
  SGB=gsub('_marker/','',strsplit(folder_list_all[[i]],split='SGB')[[1]][2])
  SGB=paste0("SGB",SGB)
  SGB=gsub('_group','',SGB)
  SGB=gsub('10068','10068s',SGB)
  taxa=taxa_list_summary$cluster_taxa[taxa_list_summary$cluster==SGB]
  cutoff=df_strainphlan_cutoff$cutoff[df_strainphlan_cutoff$X==taxa]
  sweep=sweep_list_all$scaffold[grep(SGB,sweep_list_all$scaffold)]
  total_sweep=lapply(sweep, function(x) read.table(paste0("../../instrain/compare_select_lists/",x,"_sweep.txt"),header = F))
  names(total_sweep)=paste0('sweep',seq_along(total_sweep))
  
  if (length(sweep)>0){
    tree_file=tree_list_all[[i]]
    t=read_tree(file=tree_file)
    t$tip.label=gsub("BIS","-BIS",t$tip.label)
    t$tip.label=gsub("_marker","",t$tip.label)
    t=midpoint(t)
    t$node.label=1:t$Nnode+Ntip(t)
    newtotalstrain_list[[i]]=data.frame(V1=t$tip.label)
    names(newtotalstrain_list)[i]=taxa
    dist_all=c(get_all_pairwise_distances(t,only_clades=t$tip.label))
    if(length(cutoff)==0){
      cutoff=10^-3/median(dist_all)
    }
    t$edge.length=t$edge.length/median(dist_all)
    ratio_df=sweep_finder(t,cutoff)
    ratio_df_5=ratio_df %>% filter(ratio>=5)
   ratio_df_5$new_mean=sapply(ratio_df_5$node, function(x) get_new_mean(x))
    ratio_df_5=ratio_df_5 %>% mutate(new_ratio=case_when(is.na(new_mean) ~ ratio, TRUE ~ mean_outdist/new_mean))             
   ratio_df_5f= ratio_df_5 %>% filter(new_ratio>=5)   
    newsweep=lapply(ratio_df_5f$node, function(x) {y=as.data.frame(get_subtree_at_node(t, x-Ntip(t))$subtree$tip.label);
    names(y)="V1";y})
    newsweep=Filter(function(x) nrow(x) > 2, newsweep)
    
    
    if (length(newsweep>0)){
      names(newsweep)=paste0('newsweep',seq_along(newsweep))
      newtotalsweep=merge_sweeps(newsweep,list())
      #newtotalsweep=newsweep
      
      rf_list[[i]]=c(compare_sweeps_2_withMG(total_sweep,newtotalsweep))
      sf_list[[i]]=c(length(compare_sweeps_withMG(total_sweep,newtotalsweep))/length(total_sweep),
                     length(compare_sweeps_withMG(newtotalsweep,total_sweep))/length(newtotalsweep))
      newtotalsweep_list[[i]]=newtotalsweep
      names(rf_list)[i]=taxa
      names(sf_list)[i]=taxa
      names(newtotalsweep_list)[i]=taxa
    }
  }
}



```



## import metadata
```{r metagenomes}

baseline_data_age=read.csv('baseline_data_age.csv')
baseline_data_CRC=read.csv('baseline_data_CRC.csv')
baseline_data_IBD=read.csv('baseline_data_IBD.csv')
baseline_data_UC=read.csv('baseline_data_UC.csv')
baseline_data_CD=read.csv('baseline_data_CD.csv')
baseline_data_T2D=read.csv('baseline_data_T2D.csv')

```


```{r others}
metadata_isolate=read.csv('../../UHGG_plus4_metadata.csv') 

newtotalsweep_merged_list=lapply(newtotalsweep_list,function(x) bind_rows(x,.id='id'))
newtotalsweep_merged_list=Filter(function(x) nrow(x) > 0, newtotalsweep_merged_list)

SGB_class=read.csv('../../SGBdiffH_500_1pois_1nb_finda/SGB_classify_list_sweepy.csv')
newtotalsweep_list_c=newtotalsweep_list[!(names(newtotalsweep_list) %in% c(SGB_class$pathogen,SGB_class$probiotics))]
sum(sapply(newtotalsweep_list_c,length))
```

## age metadata sweep merge
```{r}
#in this step all isolates are removed, since we only merge with MG metadata,
#this is for age
newtotalsweep_merged_list_1=lapply(newtotalsweep_merged_list,function(x)
  merge(baseline_data_age,x,by.x='sample_id',by.y='V1') %>% select(sample_id,subject_id,id,disease,disease_subtype,age,age_category,country,family,non_westernized,antibiotics_current_use))

#this step makes family either family or subject id, so then if we slice for family,
#we end up slicing for both subject and family
newtotalsweep_merged_list_1=lapply(newtotalsweep_merged_list_1,function(x)
  x %>% mutate(family= ifelse(is.na(family),subject_id, family)) %>%
#    filter(!(age_category %in% c('newborn','child','schoolage'))) %>%
#    filter(disease=='healthy') %>%    #test what happens if I only consider old age and no disease
#    filter(antibiotics_current_use=='no') %>%
    group_by(id,family) %>% slice(1) %>% ungroup() %>% group_by(id) %>% 
    filter(n()>2) %>% ungroup())

newtotalsweep_merged_list_1=Filter(function(x) nrow(x) > 0, newtotalsweep_merged_list_1)

newtotalsweep_merged_list_1c=newtotalsweep_merged_list_1[!(names(newtotalsweep_merged_list_1) %in% c(SGB_class$pathogen,SGB_class$probiotics))]
sum(sapply(newtotalsweep_merged_list_1c,function(x)length(unique(x$id))))

newtotalstrain_list_1=newtotalstrain_list[names(newtotalsweep_merged_list_1)]

newtotalstrain_list_1=lapply(newtotalstrain_list_1,function(x)
  merge(baseline_data_age,x,by.x='sample_id',by.y='V1') %>% select(sample_id,subject_id,disease,disease_subtype,age,age_category,country,family,non_westernized,antibiotics_current_use)%>% mutate(age_category = ifelse(is.na(age_category), "adult", age_category)))


newtotalstrain_list_1=mapply(function (x,y) x %>%
  group_by(family) %>%
  group_modify(~ {
    if (any(.x$sample_id %in% y$sample_id)) {
      .x %>% filter(sample_id %in% y$sample_id) %>% slice_sample(n = 1)
    } else {
      .x %>% slice_sample(n = 1)
    }
  }) %>%
  ungroup(),
  newtotalstrain_list_1,
  newtotalsweep_merged_list_1,
  SIMPLIFY = F
)
```

##age sweep association
```{r,warning=FALSE}
# age
newtotalsweep_unmerged_list_1=lapply(newtotalsweep_merged_list_1,
                                     function(x) split(x,f=x$id))
new_issweep_list_unmerged_1=mapply(function(x,y) 
  lapply(x,function(z) y=y %>% mutate(issweep=case_when(sample_id %in% z$sample_id ~ 'sweep', TRUE~ 'notsweep'))),
  newtotalsweep_unmerged_list_1,newtotalstrain_list_1,SIMPLIFY = F)


new_issweep_p_unmerged_1=list()
new_issweep_R2_unmerged_1=list()

for (i in 1:length(new_issweep_list_unmerged_1)){
  new_issweep_list_unmerged_1s=lapply(new_issweep_list_unmerged_1[[i]], function(x) x%>% select(age_category, issweep))
  new_issweep_list_unmerged_1s=do.call(cbind,new_issweep_list_unmerged_1s)
  issweep_names=grepl('issweep',names(new_issweep_list_unmerged_1s))
  new_issweep_list_unmerged_1s=cbind(new_issweep_list_unmerged_1s[,1],
                                     new_issweep_list_unmerged_1s[issweep_names])
  names(new_issweep_list_unmerged_1s)[1]='age_category'
  new_issweep_list_unmerged_1s=new_issweep_list_unmerged_1s %>% select(-where(~ all(. == "notsweep")))
  new_issweep_list_unmerged_1s=new_issweep_list_unmerged_1s %>% select(-where(~ all(. == "sweep")))
  new_issweep_list_unmerged_1s=mutate_if(new_issweep_list_unmerged_1s, is.character, as.factor)
  if (ncol(new_issweep_list_unmerged_1s)>1){
    new_issweep_list_unmerged_1s_glm=glm(age_category~.,
                                         data=new_issweep_list_unmerged_1s,family='binomial')
    
    new_issweep_list_unmerged_1s_glm_null = glm(age_category~1,
                                                data=new_issweep_list_unmerged_1s,family='binomial')
    
    scope = list(lower=formula(new_issweep_list_unmerged_1s_glm_null),
                 upper=formula(new_issweep_list_unmerged_1s_glm))
    
    select.fit = SignifReg(new_issweep_list_unmerged_1s_glm_null, scope = scope, direction = "forward", trace = FALSE,adjust.method = "fdr",alpha = 0.1)
    R2=1-(logLik(select.fit) / logLik(new_issweep_list_unmerged_1s_glm_null))
    new_issweep_R2_unmerged_1[[i]]=R2
    
    new_issweep_tab_unmerged_1=lapply(new_issweep_list_unmerged_1[[i]], function(x) x%>% count(issweep,age_category) %>% spread(age_category,n,fill = 0))
    new_issweep_tab_unmerged_1=lapply(new_issweep_tab_unmerged_1, function(x) x%>% 
                                        select(-issweep))
    
    df_psweep=as.data.frame(select.fit$coefficients)
    df_psweep$sweep=row.names(df_psweep)
    rownames(df_psweep)=NULL
    names(df_psweep)[1]='coef'
    df_psweep$p.value=select.fit$steps.info$max_pvalue
    df_psweep$sweep=gsub('.issweepsweep','',df_psweep$sweep)
    df_psweep=df_psweep %>% filter(!is.na(p.value))
    if (nrow(df_psweep)>=1){
      for (s in 1:nrow(df_psweep)){
        ss=df_psweep$sweep[s]
        df_psweep$fp.value[s]=fisher.test(new_issweep_tab_unmerged_1[ss][[1]])$p.value
      }
      df_psweep$p.adjust=p.adjust(df_psweep$fp.value,method='fdr')
    }
  }else{
    df_psweep=data.frame(coef=0,sweep=NA,p.value=NA,fp.value=NA,p.adjust=NA)
  }
  
  new_issweep_p_unmerged_1[[i]]=df_psweep
}

names(new_issweep_p_unmerged_1)=names(new_issweep_list_unmerged_1)
new_issweep_p_df_unmerged_1=bind_rows(new_issweep_p_unmerged_1,.id='SGB')
new_issweep_p_df_unmerged_1=new_issweep_p_df_unmerged_1 %>% filter(!is.na(p.value))
age_list=new_issweep_p_df_unmerged_1$p.value<0.05  #actually there has been a filter so this step is just here so the variables stay the same as before
new_issweep_p_df_unmerged_1_age=new_issweep_p_df_unmerged_1[age_list,]

#filter for commensals
SGB_class=read.csv('../../SGBdiffH_500_1pois_1nb_finda/SGB_classify_list_sweepy.csv')
new_issweep_p_df_unmerged_1_age=new_issweep_p_df_unmerged_1_age %>% 
  mutate(class=case_when(SGB %in% SGB_class$pathogen ~ 'pathogens',
                                SGB %in% SGB_class$probiotics ~ 'probiotics',
                                TRUE ~ 'commensals'))
new_issweep_p_df_unmerged_1_age=new_issweep_p_df_unmerged_1_age %>% 
  filter(class=='commensals')

new_issweep_list_unmerged_1_age=list()
new_issweep_geo_p_unmerged_1_age=list()
for (s in 1:nrow(new_issweep_p_df_unmerged_1_age)){
  taxa=new_issweep_p_df_unmerged_1_age$SGB[s]
  sweep=new_issweep_p_df_unmerged_1_age$sweep[s]
 
  stt=new_issweep_list_unmerged_1[taxa]
  sweep_list=stt[[1]][sweep][[1]] %>% filter(issweep=='sweep')
  new_issweep_list_unmerged_1_age[[s]]=sweep_list
  names(new_issweep_list_unmerged_1_age)[s]=paste(taxa,sweep,sep='-')
  # is sweep dominated by a country, and also check sweep direction (enrich for age or no)
  observed_freq <- table(sweep_list$country)
  if (length(observed_freq)>1){
    new_issweep_geo_p_unmerged_1_age[[s]]=c(chisq.test(observed_freq)$p.value,new_issweep_p_df_unmerged_1_age$coef[s]>0)
  }else{
    new_issweep_geo_p_unmerged_1_age[[s]]=c(0,new_issweep_p_df_unmerged_1_age$coef[s]>0)
  }
  names(new_issweep_geo_p_unmerged_1_age)[s]=paste(taxa,sweep,sep='-')
}

new_issweep_geo_p_unmerged_1_age_df=as.data.frame(t(bind_rows(new_issweep_geo_p_unmerged_1_age,.id='id')))
names(new_issweep_geo_p_unmerged_1_age_df)=c('geo_enriched','age_enriched')
new_issweep_geo_p_unmerged_1_age_df=new_issweep_geo_p_unmerged_1_age_df %>%
  mutate(geo_enriched=case_when(geo_enriched<=0.05 ~ TRUE, TRUE ~ FALSE))
new_issweep_geo_p_unmerged_1_age_df=new_issweep_geo_p_unmerged_1_age_df %>%
  mutate(age_enriched=case_when(age_enriched=='1' ~ 'enriched', TRUE ~ 'depleted'))

new_issweep_geo_p_unmerged_1_age_df_nogeo=new_issweep_geo_p_unmerged_1_age_df %>%
  filter(geo_enriched==FALSE) %>% select(age_enriched)
names(new_issweep_geo_p_unmerged_1_age_df_nogeo)='age_sweep_NONGEO'


```


```{r}
colors=c('#4393C3','#D6604D')
p1_age=Heatmap(new_issweep_geo_p_unmerged_1_age_df_nogeo,col=colors,row_names_max_width = unit(12, "cm"))
png('non_geo_sweeps_reg_age_cleaned_0.05.png',width=2000,height=2500,res=300)
print(p1_age)
dev.off()
```


#CRC metadata sweep merge
```{r}

newtotalsweep_merged_list_2=lapply(newtotalsweep_merged_list,function(x)
  merge(baseline_data_CRC,x,by.x='sample_id',by.y='V1') %>% select(sample_id,subject_id,id,disease,age,age_category,country,family,non_westernized,antibiotics_current_use))
# %>%   filter(antibiotics_current_use=='no')
#  %>%  mutate(disease = case_when(grepl("CRC", disease, ignore.case = TRUE) ~ "CRC",
#                                 grepl("carcinoma", disease, ignore.case = TRUE) ~ "CRC",
#                                 TRUE ~ 'non-CRC')))

newtotalsweep_merged_list_2=lapply(newtotalsweep_merged_list_2,function(x)
  x %>% 
#  mutate(family= ifelse(is.na(family),subject_id, family)) %>%
#   filter(!(age_category %in% c('newborn','child','schoolage'))) %>%
        group_by(id,family) %>% slice(1) %>% ungroup() %>% 
    group_by(id) %>% 
    filter(n()>2) %>% 
  ungroup())

newtotalsweep_merged_list_2=Filter(function(x) nrow(x) > 0, newtotalsweep_merged_list_2)


newtotalstrain_list_2=newtotalstrain_list[names(newtotalsweep_merged_list_2)]

newtotalstrain_list_2=lapply(newtotalstrain_list_2,function(x)
  merge(baseline_data_CRC,x,by.x='sample_id',by.y='V1') %>% select(sample_id,subject_id,disease,disease_subtype,age,age_category,country,family, non_westernized,antibiotics_current_use))
# %>% filter(antibiotics_current_use=='no') %>%
#  mutate(disease = case_when(grepl("CRC", disease, ignore.case = TRUE) ~ "CRC",
#                                 grepl("carcinoma", disease, ignore.case = TRUE) ~ "CRC",
#                                 TRUE ~ 'non-CRC')))

newtotalstrain_list_2=mapply(function (x,y) x %>%
  group_by(family) %>%
  group_modify(~ {
    if (any(.x$sample_id %in% y$sample_id)) {
      .x %>% filter(sample_id %in% y$sample_id) %>% slice_sample(n = 1)
    } else {
      .x %>% slice_sample(n = 1)
    }
  }) %>%
  ungroup(),
  newtotalstrain_list_2,
  newtotalsweep_merged_list_2,
  SIMPLIFY = F
)

```

#CRC sweep association
```{r,echo=FALSE,warning=FALSE}
# CRC
newtotalsweep_unmerged_list_2=lapply(newtotalsweep_merged_list_2,
                                     function(x) split(x,f=x$id))
new_issweep_list_unmerged_2=mapply(function(x,y) 
  lapply(x,function(z) y=y %>% mutate(issweep=case_when(sample_id %in% z$sample_id ~ 'sweep', TRUE~ 'notsweep'))),
  newtotalsweep_unmerged_list_2,newtotalstrain_list_2,SIMPLIFY = F)


new_issweep_p_unmerged_2=list()
new_issweep_R2_unmerged_2=list()

#i=17 (for B.frag)
#i=15 for B.intestinalis
i=15
i=21
for (i in 1:length(new_issweep_list_unmerged_2)){
  new_issweep_list_unmerged_2s=lapply(new_issweep_list_unmerged_2[[i]], function(x) x%>% select(disease, issweep))
  new_issweep_list_unmerged_2s=do.call(cbind,new_issweep_list_unmerged_2s)
  issweep_names=grepl('issweep',names(new_issweep_list_unmerged_2s))
  new_issweep_list_unmerged_2s=cbind(new_issweep_list_unmerged_2s[,1],
                                     new_issweep_list_unmerged_2s[issweep_names])
  names(new_issweep_list_unmerged_2s)[1]='disease'
  new_issweep_list_unmerged_2s=new_issweep_list_unmerged_2s %>% select(-where(~ all(. == "notsweep")))
  new_issweep_list_unmerged_2s=new_issweep_list_unmerged_2s %>% select(-where(~ all(. == "sweep")))
  new_issweep_list_unmerged_2s=mutate_if(new_issweep_list_unmerged_2s, is.character, as.factor)
  
  if (ncol(new_issweep_list_unmerged_2s)>1) {

#    # Create model matrix (without response variable)
  #  library(glmnet)
  #    X <- model.matrix(~ ., data = new_issweep_list_unmerged_2s)[, -1]  # Remove intercept
  #    cvfit <- cv.glmnet(X[,-1], X[,1], alpha = 1, family = "binomial")

      # Find linearly dependent columns
#      lc <- findLinearCombos(X)
      
#      if (length(lc$remove)>0){
        # Remove redundant predictors
#      new_issweep_list_unmerged_2s <- new_issweep_list_unmerged_2s[, -lc$remove]
#      }
    new_issweep_list_unmerged_2s_glm=glm(disease~.,
                                         data=new_issweep_list_unmerged_2s,family='binomial')
    
    new_issweep_list_unmerged_2s_glm_null = glm(disease~1,
                                                data=new_issweep_list_unmerged_2s,family='binomial')
    
    scope = list(lower=formula(new_issweep_list_unmerged_2s_glm_null),
                 upper=formula(new_issweep_list_unmerged_2s_glm))
    
    select.fit = SignifReg(new_issweep_list_unmerged_2s_glm_null, scope = scope, direction = "forward", trace = FALSE,adjust.method = "fdr",alpha=0.1)
    R2=1-(logLik(select.fit) / logLik(new_issweep_list_unmerged_2s_glm_null))
    new_issweep_R2_unmerged_2[[i]]=R2
    
    new_issweep_tab_unmerged_2=lapply(new_issweep_list_unmerged_2[[i]], function(x) x%>% count(issweep,disease) %>% spread(disease,n,fill = 0))
    new_issweep_tab_unmerged_2=lapply(new_issweep_tab_unmerged_2, function(x) x%>% 
                                        select(-issweep))
    
    df_psweep=as.data.frame(select.fit$coefficients)
    df_psweep$sweep=row.names(df_psweep)
    rownames(df_psweep)=NULL
    names(df_psweep)[1]='coef'
    df_psweep$p.value=select.fit$steps.info$max_pvalue
    df_psweep$sweep=gsub('.issweepsweep','',df_psweep$sweep)
    df_psweep=df_psweep %>% filter(!is.na(p.value))
    if (nrow(df_psweep)>=1){
      for (s in 1:nrow(df_psweep)){
        ss=df_psweep$sweep[s]
        df_psweep$fp.value[s]=fisher.test(new_issweep_tab_unmerged_2[ss][[1]])$p.value
      }
      df_psweep$p.adjust=p.adjust(df_psweep$fp.value,method='fdr')
    }
  }else{
    df_psweep=data.frame(coef=0,sweep=NA,p.value=NA,fp.value=NA,p.adjust=NA)
  }
  
  new_issweep_p_unmerged_2[[i]]=df_psweep
}

names(new_issweep_p_unmerged_2)=names(new_issweep_list_unmerged_2)
new_issweep_p_df_unmerged_2=bind_rows(new_issweep_p_unmerged_2,.id='SGB')
new_issweep_p_df_unmerged_2=new_issweep_p_df_unmerged_2 %>% filter(!is.na(p.value))

CRC_list=new_issweep_p_df_unmerged_2$p.value<0.05
new_issweep_p_df_unmerged_2_CRC=new_issweep_p_df_unmerged_2[CRC_list,]


ggplot(data = new_issweep_p_df_unmerged_2, aes(x = coef, y = -log10(p.value))) +
         geom_point()+geom_hline(yintercept = 2)

#new_issweep_p_df_unmerged_2_CRC_list=split(new_issweep_p_df_unmerged_2_CRC,new_issweep_p_df_unmerged_2_CRC$SGB)
#new_issweep_p_df_unmerged_2_CRC_list_1=list()
#for (s in 1:length(new_issweep_p_df_unmerged_2_CRC_list)){
#  taxa=names(new_issweep_p_df_unmerged_2_CRC_list)[s]
#  sweeps=new_issweep_p_df_unmerged_2_CRC_list[[taxa]]$sweep
#  sweep_list=lapply(sweeps,function(x) data.frame(V1=newtotalsweep_unmerged_list_2[[taxa]][[x]]$sample_id))
#  names(sweep_list)=sweeps
#  sweep_list=merge_sweeps(sweep_list,list())
#  new_issweep_p_df_unmerged_2_CRC_list_1[[s]]=new_issweep_p_df_unmerged_2_CRC_list[[s]]%>% filter(sweep %in% names(sweep_list))
#}
#new_issweep_p_df_unmerged_2_CRC=bind_rows(new_issweep_p_df_unmerged_2_CRC_list_1)

#filter for commensals
new_issweep_p_df_unmerged_2_CRC=new_issweep_p_df_unmerged_2_CRC %>% 
  mutate(class=case_when(SGB %in% SGB_class$pathogen ~ 'pathogens',
                         SGB %in% SGB_class$probiotics ~ 'probiotics',
                         TRUE ~ 'commensals'))
new_issweep_p_df_unmerged_2_CRC=new_issweep_p_df_unmerged_2_CRC %>% 
  filter(class=='commensals')


new_issweep_list_unmerged_2_CRC=list()
new_issweep_geo_p_unmerged_2_CRC=list()
for (s in 1:nrow(new_issweep_p_df_unmerged_2_CRC)){
  taxa=new_issweep_p_df_unmerged_2_CRC$SGB[s]
  sweep=new_issweep_p_df_unmerged_2_CRC$sweep[s]
  
  stt=new_issweep_list_unmerged_2[taxa]
  sweep_list=stt[[1]][sweep][[1]] %>% filter(issweep=='sweep')
  new_issweep_list_unmerged_2_CRC[[s]]=sweep_list
  names(new_issweep_list_unmerged_2_CRC)[s]=paste(taxa,sweep,sep='-')
  # is sweep dominated by a country, and also check sweep direction (enrich for CRC or no)
  observed_freq <- table(sweep_list$country)
  if (length(observed_freq)>1){
    new_issweep_geo_p_unmerged_2_CRC[[s]]=c(chisq.test(observed_freq)$p.value,new_issweep_p_df_unmerged_2_CRC$coef[s]<0)  #here needs to be <0 because the correlation is for non-CRC
  }else{
    new_issweep_geo_p_unmerged_2_CRC[[s]]=c(0,new_issweep_p_df_unmerged_2_CRC$coef[s]<0) #here needs to be <0 because the correlation is for non-CRC
  }
  names(new_issweep_geo_p_unmerged_2_CRC)[s]=paste(taxa,sweep,sep='-')
}

new_issweep_geo_p_unmerged_2_CRC_df=as.data.frame(t(bind_rows(new_issweep_geo_p_unmerged_2_CRC,.id='id')))
names(new_issweep_geo_p_unmerged_2_CRC_df)=c('geo_enriched','CRC_enriched')
new_issweep_geo_p_unmerged_2_CRC_df=new_issweep_geo_p_unmerged_2_CRC_df %>%
  mutate(geo_enriched=case_when(geo_enriched<=0.05 ~ TRUE, TRUE ~ FALSE))
new_issweep_geo_p_unmerged_2_CRC_df=new_issweep_geo_p_unmerged_2_CRC_df %>%
  mutate(CRC_enriched=case_when(CRC_enriched=='1' ~ 'enriched', TRUE ~ 'depleted'))

new_issweep_geo_p_unmerged_2_CRC_df_nogeo=new_issweep_geo_p_unmerged_2_CRC_df %>%
  filter(geo_enriched==FALSE) %>% select(CRC_enriched)
names(new_issweep_geo_p_unmerged_2_CRC_df_nogeo)='CRC_sweep_NONGEO'
write.csv(newtotalsweep_merged_list_2$s__Bacteroides_caccae_SGB1877 %>% count(id),'bc_dist_new.csv')

```

```{r}
colors=c('#4393C3','#D6604D')
p1_CRC=Heatmap(new_issweep_geo_p_unmerged_2_CRC_df_nogeo,col=colors,row_names_max_width = unit(12, "cm"))
png('non_geo_sweeps_reg_CRC_vs_non_cleaned_0.05.png',width=1800,height=3000,res=300)
print(p1_CRC)
dev.off()
```


#CRC age merge
```{r}
new_issweep_geo_p_unmerged_comb=merge(new_issweep_geo_p_unmerged_1_age_df_nogeo,
                                                      new_issweep_geo_p_unmerged_2_CRC_df_nogeo,
                                                      by=0,all=T)
new_issweep_geo_p_unmerged_comb$SGB=sapply(strsplit(new_issweep_geo_p_unmerged_comb$Row.names,split = '-'),'[[',1)

new_issweep_geo_p_unmerged_comb_dircount=new_issweep_geo_p_unmerged_comb %>% group_by(SGB) %>% 
  summarize(unique_age = n_distinct(age_sweep_NONGEO,na.rm=T),unique_CRC=n_distinct(CRC_sweep_NONGEO,na.rm=T))

oppo_SGB=new_issweep_geo_p_unmerged_comb_dircount %>% filter(unique_age>1|unique_CRC>1)
new_issweep_geo_p_unmerged_comb_filter=new_issweep_geo_p_unmerged_comb %>% filter(SGB %in% oppo_SGB$SGB) 
new_issweep_geo_p_unmerged_comb_filter$Row.names=rownames(new_issweep_geo_p_unmerged_comb_filter)
new_issweep_geo_p_unmerged_comb_filter$Row.names=NULL
names(new_issweep_geo_p_unmerged_comb_filter)=c('age','CRC','SGB')
new_issweep_geo_p_unmerged_comb_filter_1=new_issweep_geo_p_unmerged_comb_filter[,c(1:2)]


colors2=c('#4393C3','#D6604D','white')
col_mapping <- c("enriched" = '#D6604D', "depleted" = "#4393C3")
p2=Heatmap(new_issweep_geo_p_unmerged_comb_filter_1,col=col_mapping,na_col = "white",
           column_gap = unit(5, "mm"),
           column_split=factor(c('age','CRC')),
           rect_gp = gpar(col = "grey60", lwd = 1),
           row_split = factor(new_issweep_geo_p_unmerged_comb_filter$SGB),
           show_row_names = F,
           row_title_rot = 0)

png('non_geo_sweeps_reg_CRC_age_bidirectional_cleaned_0.05.png',width=1800,height=2000,res=300)
print(p2)
dev.off()

```
#IBD metadata sweep merge
```{r}
#in this step all isolates are removed, since we only merge with MG metadata,

newtotalsweep_merged_list_3=lapply(newtotalsweep_merged_list,function(x)
  merge(baseline_data_IBD,x,by.x='sample_id',by.y='V1') %>% select(sample_id,subject_id,id,disease,age,age_category,country,family,non_westernized,antibiotics_current_use))

newtotalsweep_merged_list_3=lapply(newtotalsweep_merged_list_3,function(x)
  x %>% 
#    mutate(family= ifelse(is.na(family),subject_id, family)) %>%
#    filter(!(age_category %in% c('newborn','child','schoolage'))) %>%
    group_by(id,family) %>% slice(1) %>% ungroup() %>% group_by(id) %>% 
    filter(n()>2) %>% ungroup())

newtotalsweep_merged_list_3=Filter(function(x) nrow(x) > 0, newtotalsweep_merged_list_3)

newtotalstrain_list_3=newtotalstrain_list[names(newtotalsweep_merged_list_3)]

newtotalstrain_list_3=lapply(newtotalstrain_list_3,function(x)
  merge(baseline_data_IBD,x,by.x='sample_id',by.y='V1') %>% select(sample_id,subject_id,disease,disease_subtype,age,age_category,country,family, non_westernized,antibiotics_current_use))
#  %>%filter(antibiotics_current_use=='no') %>%
#  mutate(disease=case_when(disease=='IBD'~'IBD', TRUE ~ 'non-IBD')))

newtotalstrain_list_3=mapply(function (x,y) x %>%
  group_by(family) %>%
  group_modify(~ {
    if (any(.x$sample_id %in% y$sample_id)) {
      .x %>% filter(sample_id %in% y$sample_id) %>% slice_sample(n = 1)
    } else {
      .x %>% slice_sample(n = 1)
    }
  }) %>%
  ungroup(),
  newtotalstrain_list_3,
  newtotalsweep_merged_list_3,
  SIMPLIFY = F
)
```

# IBD associations
```{r,warning=FALSE}
newtotalsweep_unmerged_list_3=lapply(newtotalsweep_merged_list_3,
                                     function(x) split(x,f=x$id))
new_issweep_list_unmerged_3=mapply(function(x,y) 
  lapply(x,function(z) y=y %>% mutate(issweep=case_when(sample_id %in% z$sample_id ~ 'sweep', TRUE~ 'notsweep'))),
  newtotalsweep_unmerged_list_3,newtotalstrain_list_3,SIMPLIFY = F)




new_issweep_p_unmerged_3=list()
new_issweep_R2_unmerged_3=list()

#i=17 (for B.frag)

for (i in 1:length(new_issweep_list_unmerged_3)){
  new_issweep_list_unmerged_3s=lapply(new_issweep_list_unmerged_3[[i]], function(x) x%>% select(disease, issweep))
  new_issweep_list_unmerged_3s=do.call(cbind,new_issweep_list_unmerged_3s)
  issweep_names=grepl('issweep',names(new_issweep_list_unmerged_3s))
  new_issweep_list_unmerged_3s=cbind(new_issweep_list_unmerged_3s[,1],
                                     new_issweep_list_unmerged_3s[issweep_names])
  names(new_issweep_list_unmerged_3s)[1]='disease'
  new_issweep_list_unmerged_3s=new_issweep_list_unmerged_3s %>% select(-where(~ all(. == "notsweep")))
  new_issweep_list_unmerged_3s=new_issweep_list_unmerged_3s %>% select(-where(~ all(. == "sweep")))
  new_issweep_list_unmerged_3s=mutate_if(new_issweep_list_unmerged_3s, is.character, as.factor)
  if (ncol(new_issweep_list_unmerged_3s)>1){
    new_issweep_list_unmerged_3s_glm=glm(disease~.,
                                         data=new_issweep_list_unmerged_3s,family='binomial')
    
    new_issweep_list_unmerged_3s_glm_null = glm(disease~1,
                                                data=new_issweep_list_unmerged_3s,family='binomial')
    
    scope = list(lower=formula(new_issweep_list_unmerged_3s_glm_null),
                 upper=formula(new_issweep_list_unmerged_3s_glm))
    
    select.fit = SignifReg(new_issweep_list_unmerged_3s_glm_null, scope = scope, direction = "forward", trace = FALSE,adjust.method = "fdr", alpha=0.1)
    R2=1-(logLik(select.fit) / logLik(new_issweep_list_unmerged_3s_glm_null))
    new_issweep_R2_unmerged_3[[i]]=R2
    
    new_issweep_tab_unmerged_3=lapply(new_issweep_list_unmerged_3[[i]], function(x) x%>% count(issweep,disease) %>% spread(disease,n,fill = 0))
    new_issweep_tab_unmerged_3=lapply(new_issweep_tab_unmerged_3, function(x) x%>% 
                                        select(-issweep))
    
    df_psweep=as.data.frame(select.fit$coefficients)
    df_psweep$sweep=row.names(df_psweep)
    rownames(df_psweep)=NULL
    names(df_psweep)[1]='coef'
    df_psweep$p.value=select.fit$steps.info$max_pvalue
    df_psweep$sweep=gsub('.issweepsweep','',df_psweep$sweep)
    df_psweep=df_psweep %>% filter(!is.na(p.value))
    if (nrow(df_psweep)>=1){
      for (s in 1:nrow(df_psweep)){
        ss=df_psweep$sweep[s]
        df_psweep$fp.value[s]=fisher.test(new_issweep_tab_unmerged_3[ss][[1]])$p.value
      }
      df_psweep$p.adjust=p.adjust(df_psweep$fp.value,method='fdr')
    }
  }else{
    df_psweep=data.frame(coef=0,sweep=NA,p.value=NA,fp.value=NA,p.adjust=NA)
  }
  
  new_issweep_p_unmerged_3[[i]]=df_psweep
}

names(new_issweep_p_unmerged_3)=names(new_issweep_list_unmerged_3)
new_issweep_p_df_unmerged_3=bind_rows(new_issweep_p_unmerged_3,.id='SGB')
new_issweep_p_df_unmerged_3=new_issweep_p_df_unmerged_3 %>% filter(!is.na(p.value))
IBD_list=new_issweep_p_df_unmerged_3$p.value<0.05
new_issweep_p_df_unmerged_3_IBD=new_issweep_p_df_unmerged_3[IBD_list,]

#filter for commensals
new_issweep_p_df_unmerged_3_IBD=new_issweep_p_df_unmerged_3_IBD %>% 
  mutate(class=case_when(SGB %in% SGB_class$pathogen ~ 'pathogens',
                         SGB %in% SGB_class$probiotics ~ 'probiotics',
                         TRUE ~ 'commensals'))
new_issweep_p_df_unmerged_3_IBD=new_issweep_p_df_unmerged_3_IBD %>% 
  filter(class=='commensals')


new_issweep_list_unmerged_3_IBD=list()
new_issweep_geo_p_unmerged_3_IBD=list()
for (s in 1:nrow(new_issweep_p_df_unmerged_3_IBD)){
  taxa=new_issweep_p_df_unmerged_3_IBD$SGB[s]
  sweep=new_issweep_p_df_unmerged_3_IBD$sweep[s]
  
  stt=new_issweep_list_unmerged_3[taxa]
  sweep_list=stt[[1]][sweep][[1]] %>% filter(issweep=='sweep')
  new_issweep_list_unmerged_3_IBD[[s]]=sweep_list
  names(new_issweep_list_unmerged_3_IBD)[s]=paste(taxa,sweep,sep='-')
  # is sweep dominated by a country, and also check sweep direction (enrich for IBD or no)
  observed_freq <- table(sweep_list$country)
  if (length(observed_freq)>1){
    new_issweep_geo_p_unmerged_3_IBD[[s]]=c(chisq.test(observed_freq)$p.value,new_issweep_p_df_unmerged_3_IBD$coef[s]<0)  #here needs to be <0 because the correlation is for non-IBD
  }else{
    new_issweep_geo_p_unmerged_3_IBD[[s]]=c(0,new_issweep_p_df_unmerged_3_IBD$coef[s]<0) 
  }
  names(new_issweep_geo_p_unmerged_3_IBD)[s]=paste(taxa,sweep,sep='-')
}

new_issweep_geo_p_unmerged_3_IBD_df=as.data.frame(t(bind_rows(new_issweep_geo_p_unmerged_3_IBD,.id='id')))
names(new_issweep_geo_p_unmerged_3_IBD_df)=c('geo_enriched','IBD_enriched')
new_issweep_geo_p_unmerged_3_IBD_df=new_issweep_geo_p_unmerged_3_IBD_df %>%
  mutate(geo_enriched=case_when(geo_enriched<0.05 ~ TRUE, TRUE ~ FALSE))
new_issweep_geo_p_unmerged_3_IBD_df=new_issweep_geo_p_unmerged_3_IBD_df %>%
  mutate(IBD_enriched=case_when(IBD_enriched=='1' ~ 'enriched', TRUE ~ 'depleted'))

new_issweep_geo_p_unmerged_3_IBD_df_nogeo=new_issweep_geo_p_unmerged_3_IBD_df %>%
  filter(geo_enriched==FALSE) %>% select(IBD_enriched)
names(new_issweep_geo_p_unmerged_3_IBD_df_nogeo)='IBD_sweep_NONGEO'


```

```{r}
colors=c('#4393C3','#D6604D')
p1_IBD=Heatmap(new_issweep_geo_p_unmerged_3_IBD_df_nogeo,col=colors,row_names_max_width = unit(12, "cm"))
png('non_geo_sweeps_reg_IBD_vs_non_clean_0.05.png',width=2000,height=4500,res=300)
print(p1_IBD)
dev.off()
```

#CD metadata sweep merge
```{r}
#in this step all isolates are removed, since we only merge with MG metadata,
#this is for CD
newtotalsweep_merged_list_3a=lapply(newtotalsweep_merged_list,function(x)
  merge(baseline_data_CD,x,by.x='sample_id',by.y='V1') %>% select(sample_id,subject_id,id,disease,disease_subtype,age,age_category,country,family,non_westernized,antibiotics_current_use))

newtotalsweep_merged_list_3a=lapply(newtotalsweep_merged_list_3a,function(x)
  x %>% 
 #   mutate(family= ifelse(is.na(family),subject_id, family)) %>%
 #   filter(!(age_category %in% c('newborn','child','schoolage'))) %>%
    group_by(id,family) %>% slice(1) %>% ungroup() %>% group_by(id) %>% 
    filter(n()>2) %>% ungroup())

newtotalsweep_merged_list_3a=Filter(function(x) nrow(x) > 0, newtotalsweep_merged_list_3a)


newtotalstrain_list_3a=newtotalstrain_list[names(newtotalsweep_merged_list_3a)]

newtotalstrain_list_3a=lapply(newtotalstrain_list_3a,function(x)
  merge(baseline_data_CD,x,by.x='sample_id',by.y='V1') %>% select(sample_id,subject_id,disease,disease_subtype,age,age_category,country,family,non_westernized,antibiotics_current_use))


newtotalstrain_list_3a=mapply(function (x,y) x %>%
  group_by(family) %>%
  group_modify(~ {
    if (any(.x$sample_id %in% y$sample_id)) {
      .x %>% filter(sample_id %in% y$sample_id) %>% slice_sample(n = 1)
    } else {
      .x %>% slice_sample(n = 1)
    }
  }) %>%
  ungroup(),
  newtotalstrain_list_3a,
  newtotalsweep_merged_list_3a,
  SIMPLIFY = F
)


```


## CD associations
```{r,warning=FALSE}
# CD
newtotalsweep_unmerged_list_3a=lapply(newtotalsweep_merged_list_3a,
                                     function(x) split(x,f=x$id))
new_issweep_list_unmerged_3a=mapply(function(x,y) 
  lapply(x,function(z) y=y %>% mutate(issweep=case_when(sample_id %in% z$sample_id ~ 'sweep', TRUE~ 'notsweep'))),
  newtotalsweep_unmerged_list_3a,newtotalstrain_list_3a,SIMPLIFY = F)


new_issweep_p_unmerged_3a=list()
new_issweep_R2_unmerged_3a=list()

#i=17 (for B.frag)

for (i in 1:length(new_issweep_list_unmerged_3a)){
  new_issweep_list_unmerged_3as=lapply(new_issweep_list_unmerged_3a[[i]], function(x) x%>% select(disease, issweep))
  new_issweep_list_unmerged_3as=do.call(cbind,new_issweep_list_unmerged_3as)
  issweep_names=grepl('issweep',names(new_issweep_list_unmerged_3as))
  new_issweep_list_unmerged_3as=cbind(new_issweep_list_unmerged_3as[,1],
                                     new_issweep_list_unmerged_3as[issweep_names])
  names(new_issweep_list_unmerged_3as)[1]='disease'
  new_issweep_list_unmerged_3as=new_issweep_list_unmerged_3as %>% select(-where(~ all(. == "notsweep")))
  new_issweep_list_unmerged_3as=new_issweep_list_unmerged_3as %>% select(-where(~ all(. == "sweep")))
  new_issweep_list_unmerged_3as=mutate_if(new_issweep_list_unmerged_3as, is.character, as.factor)
  if (ncol(new_issweep_list_unmerged_3as)>1){
    new_issweep_list_unmerged_3as_glm=glm(disease~.,
                                         data=new_issweep_list_unmerged_3as,family='binomial')
    
    new_issweep_list_unmerged_3as_glm_null = glm(disease~1,
                                                data=new_issweep_list_unmerged_3as,family='binomial')
    
    scope = list(lower=formula(new_issweep_list_unmerged_3as_glm_null),
                 upper=formula(new_issweep_list_unmerged_3as_glm))
    
    select.fit = SignifReg(new_issweep_list_unmerged_3as_glm_null, scope = scope, direction = "forward", trace = FALSE,adjust.method = "fdr",alpha=0.1)
    R2=1-(logLik(select.fit) / logLik(new_issweep_list_unmerged_3as_glm_null))
    new_issweep_R2_unmerged_3a[[i]]=R2
    
    new_issweep_tab_unmerged_3a=lapply(new_issweep_list_unmerged_3a[[i]], function(x) x%>% count(issweep,disease) %>% spread(disease,n,fill = 0))
    new_issweep_tab_unmerged_3a=lapply(new_issweep_tab_unmerged_3a, function(x) x%>% 
                                        select(-issweep))
    
    df_psweep=as.data.frame(select.fit$coefficients)
    df_psweep$sweep=row.names(df_psweep)
    rownames(df_psweep)=NULL
    names(df_psweep)[1]='coef'
    df_psweep$p.value=select.fit$steps.info$max_pvalue
    df_psweep$sweep=gsub('.issweepsweep','',df_psweep$sweep)
    df_psweep=df_psweep %>% filter(!is.na(p.value))
    if (nrow(df_psweep)>=1){
      for (s in 1:nrow(df_psweep)){
        ss=df_psweep$sweep[s]
        df_psweep$fp.value[s]=fisher.test(new_issweep_tab_unmerged_3a[ss][[1]])$p.value
      }
      df_psweep$p.adjust=p.adjust(df_psweep$fp.value,method='fdr')
    }
  }else{
    df_psweep=data.frame(coef=0,sweep=NA,p.value=NA,fp.value=NA,p.adjust=NA)
  }
  
  new_issweep_p_unmerged_3a[[i]]=df_psweep
}

names(new_issweep_p_unmerged_3a)=names(new_issweep_list_unmerged_3a)
new_issweep_p_df_unmerged_3a=bind_rows(new_issweep_p_unmerged_3a,.id='SGB')
new_issweep_p_df_unmerged_3a=new_issweep_p_df_unmerged_3a %>% filter(!is.na(p.value))
CD_list=new_issweep_p_df_unmerged_3a$p.value<0.05
new_issweep_p_df_unmerged_3a_CD=new_issweep_p_df_unmerged_3a[CD_list,]

#filter for commensals
new_issweep_p_df_unmerged_3a_CD=new_issweep_p_df_unmerged_3a_CD %>% 
  mutate(class=case_when(SGB %in% SGB_class$pathogen ~ 'pathogens',
                         SGB %in% SGB_class$probiotics ~ 'probiotics',
                         TRUE ~ 'commensals'))
new_issweep_p_df_unmerged_3a_CD=new_issweep_p_df_unmerged_3a_CD %>% 
  filter(class=='commensals')


new_issweep_list_unmerged_3a_CD=list()
new_issweep_geo_p_unmerged_3a_CD=list()
for (s in 1:nrow(new_issweep_p_df_unmerged_3a_CD)){
  taxa=new_issweep_p_df_unmerged_3a_CD$SGB[s]
  sweep=new_issweep_p_df_unmerged_3a_CD$sweep[s]
  
  stt=new_issweep_list_unmerged_3a[taxa]
  sweep_list=stt[[1]][sweep][[1]] %>% filter(issweep=='sweep')
  new_issweep_list_unmerged_3a_CD[[s]]=sweep_list
  names(new_issweep_list_unmerged_3a_CD)[s]=paste(taxa,sweep,sep='-')
  # is sweep dominated by a country, and also check sweep direction (enrich for CD or no)
  observed_freq <- table(sweep_list$country)
  if (length(observed_freq)>1){
    new_issweep_geo_p_unmerged_3a_CD[[s]]=c(chisq.test(observed_freq)$p.value,new_issweep_p_df_unmerged_3a_CD$coef[s]<0)  #here needs to be <0 because the correlation is for non-CD
  }else{
    new_issweep_geo_p_unmerged_3a_CD[[s]]=c(0,new_issweep_p_df_unmerged_3a_CD$coef[s]<0) #here needs to be <0 because the correlation is for non-CD
  }
  names(new_issweep_geo_p_unmerged_3a_CD)[s]=paste(taxa,sweep,sep='-')
}

new_issweep_geo_p_unmerged_3a_CD_df=as.data.frame(t(bind_rows(new_issweep_geo_p_unmerged_3a_CD,.id='id')))
names(new_issweep_geo_p_unmerged_3a_CD_df)=c('geo_enriched','CD_enriched')
new_issweep_geo_p_unmerged_3a_CD_df=new_issweep_geo_p_unmerged_3a_CD_df %>%
  mutate(geo_enriched=case_when(geo_enriched<=0.05 ~ TRUE, TRUE ~ FALSE))
new_issweep_geo_p_unmerged_3a_CD_df=new_issweep_geo_p_unmerged_3a_CD_df %>%
  mutate(CD_enriched=case_when(CD_enriched=='1' ~ 'enriched', TRUE ~ 'depleted'))

new_issweep_geo_p_unmerged_3a_CD_df_nogeo=new_issweep_geo_p_unmerged_3a_CD_df %>%
  filter(geo_enriched==FALSE) %>% select(CD_enriched)
names(new_issweep_geo_p_unmerged_3a_CD_df_nogeo)='CD_sweep_NONGEO'


```

```{r}
colors=c('#4393C3','#D6604D')
p1_CD=Heatmap(new_issweep_geo_p_unmerged_3a_CD_df_nogeo,col=colors,row_names_max_width = unit(12, "cm"))
png('non_geo_sweeps_reg_CD_vs_non_clean_0.05.png',width=2000,height=5000,res=300)
print(p1_CD)
dev.off()


```
## UC metadata sweep merge

```{r}
#this is for UC

newtotalsweep_merged_list_3b=lapply(newtotalsweep_merged_list,function(x)
  merge(baseline_data_UC,x,by.x='sample_id',by.y='V1') %>% select(sample_id,subject_id,id,disease,disease_subtype,age,age_category,country,family,non_westernized,antibiotics_current_use))

newtotalsweep_merged_list_3b=lapply(newtotalsweep_merged_list_3b,function(x)
  x %>% 
#    mutate(family= ifelse(is.na(family),subject_id, family)) %>%
#    filter(!(age_category %in% c('newborn','child','schoolage'))) %>%
    group_by(id,family) %>% slice(1) %>% ungroup() %>% group_by(id) %>% 
    filter(n()>2) %>% ungroup())

newtotalsweep_merged_list_3b=Filter(function(x) nrow(x) > 0, newtotalsweep_merged_list_3b)

newtotalstrain_list_3b=newtotalstrain_list[names(newtotalsweep_merged_list_3b)]

newtotalstrain_list_3b=lapply(newtotalstrain_list_3b,function(x)
  merge(baseline_data_UC,x,by.x='sample_id',by.y='V1') %>% select(sample_id,subject_id,disease,disease_subtype,age,age_category,country,family,non_westernized,antibiotics_current_use)) 
# %>%  filter(antibiotics_current_use=='no') %>%
#    mutate(disease=case_when(disease_subtype=='CD'~'CD',disease_subtype=='UC'~'UC', TRUE ~ 'non-UC')) %>%
#    filter(disease!='CD'))

newtotalstrain_list_3b=mapply(function (x,y) x %>%
  group_by(family) %>%
  group_modify(~ {
    if (any(.x$sample_id %in% y$sample_id)) {
      .x %>% filter(sample_id %in% y$sample_id) %>% slice_sample(n = 1)
    } else {
      .x %>% slice_sample(n = 1)
    }
  }) %>%
  ungroup(),
  newtotalstrain_list_3b,
  newtotalsweep_merged_list_3b,
  SIMPLIFY = F
)

```

# UC association
```{r,warning=F}
# UC
newtotalsweep_unmerged_list_3b=lapply(newtotalsweep_merged_list_3b,
                                      function(x) split(x,f=x$id))
new_issweep_list_unmerged_3b=mapply(function(x,y) 
  lapply(x,function(z) y=y %>% mutate(issweep=case_when(sample_id %in% z$sample_id ~ 'sweep', TRUE~ 'notsweep'))),
  newtotalsweep_unmerged_list_3b,newtotalstrain_list_3b,SIMPLIFY = F)

new_issweep_p_unmerged_3b=list()
new_issweep_R2_unmerged_3b=list()

#i=17 (for B.frag)

for (i in 1:length(new_issweep_list_unmerged_3b)){
  new_issweep_list_unmerged_3bs=lapply(new_issweep_list_unmerged_3b[[i]], function(x) x%>% select(disease, issweep))
  new_issweep_list_unmerged_3bs=do.call(cbind,new_issweep_list_unmerged_3bs)
  issweep_names=grepl('issweep',names(new_issweep_list_unmerged_3bs))
  new_issweep_list_unmerged_3bs=cbind(new_issweep_list_unmerged_3bs[,1],
                                      new_issweep_list_unmerged_3bs[issweep_names])
  names(new_issweep_list_unmerged_3bs)[1]='disease'
  new_issweep_list_unmerged_3bs=new_issweep_list_unmerged_3bs %>% select(-where(~ all(. == "notsweep")))
  new_issweep_list_unmerged_3bs=new_issweep_list_unmerged_3bs %>% select(-where(~ all(. == "sweep")))
  new_issweep_list_unmerged_3bs=mutate_if(new_issweep_list_unmerged_3bs, is.character, as.factor)
  if (ncol(new_issweep_list_unmerged_3bs)>1){
    new_issweep_list_unmerged_3bs_glm=glm(disease~.,
                                          data=new_issweep_list_unmerged_3bs,family='binomial')
    
    new_issweep_list_unmerged_3bs_glm_null = glm(disease~1,
                                                 data=new_issweep_list_unmerged_3bs,family='binomial')
    
    scope = list(lower=formula(new_issweep_list_unmerged_3bs_glm_null),
                 upper=formula(new_issweep_list_unmerged_3bs_glm))
    
    select.fit = SignifReg(new_issweep_list_unmerged_3bs_glm_null, scope = scope, direction = "forward", trace = FALSE,adjust.method = "fdr",alpha=0.1)
    R2=1-(logLik(select.fit) / logLik(new_issweep_list_unmerged_3bs_glm_null))
    new_issweep_R2_unmerged_3b[[i]]=R2
    
    new_issweep_tab_unmerged_3b=lapply(new_issweep_list_unmerged_3b[[i]], function(x) x%>% count(issweep,disease) %>% spread(disease,n,fill = 0))
    new_issweep_tab_unmerged_3b=lapply(new_issweep_tab_unmerged_3b, function(x) x%>% 
                                         select(-issweep))
    
    df_psweep=as.data.frame(select.fit$coefficients)
    df_psweep$sweep=row.names(df_psweep)
    rownames(df_psweep)=NULL
    names(df_psweep)[1]='coef'
    df_psweep$p.value=select.fit$steps.info$max_pvalue
    df_psweep$sweep=gsub('.issweepsweep','',df_psweep$sweep)
    df_psweep=df_psweep %>% filter(!is.na(p.value))
    if (nrow(df_psweep)>=1){
      for (s in 1:nrow(df_psweep)){
        ss=df_psweep$sweep[s]
        df_psweep$fp.value[s]=fisher.test(new_issweep_tab_unmerged_3b[ss][[1]])$p.value
      }
      df_psweep$p.adjust=p.adjust(df_psweep$fp.value,method='fdr')
    }
  }else{
    df_psweep=data.frame(coef=0,sweep=NA,p.value=NA,fp.value=NA,p.adjust=NA)
  }
  
  new_issweep_p_unmerged_3b[[i]]=df_psweep
}

names(new_issweep_p_unmerged_3b)=names(new_issweep_list_unmerged_3b)
new_issweep_p_df_unmerged_3b=bind_rows(new_issweep_p_unmerged_3b,.id='SGB')
new_issweep_p_df_unmerged_3b=new_issweep_p_df_unmerged_3b %>% filter(!is.na(p.value))
UC_list=new_issweep_p_df_unmerged_3b$p.value<0.05
new_issweep_p_df_unmerged_3b_UC=new_issweep_p_df_unmerged_3b[UC_list,]

#filter for commensals
new_issweep_p_df_unmerged_3b_UC=new_issweep_p_df_unmerged_3b_UC %>% 
  mutate(class=case_when(SGB %in% SGB_class$pathogen ~ 'pathogens',
                         SGB %in% SGB_class$probiotics ~ 'probiotics',
                         TRUE ~ 'commensals'))
new_issweep_p_df_unmerged_3b_UC=new_issweep_p_df_unmerged_3b_UC %>% 
  filter(class=='commensals')


new_issweep_list_unmerged_3b_UC=list()
new_issweep_geo_p_unmerged_3b_UC=list()
for (s in 1:nrow(new_issweep_p_df_unmerged_3b_UC)){
  taxa=new_issweep_p_df_unmerged_3b_UC$SGB[s]
  sweep=new_issweep_p_df_unmerged_3b_UC$sweep[s]
  
  stt=new_issweep_list_unmerged_3b[taxa]
  sweep_list=stt[[1]][sweep][[1]] %>% filter(issweep=='sweep')
  new_issweep_list_unmerged_3b_UC[[s]]=sweep_list
  names(new_issweep_list_unmerged_3b_UC)[s]=paste(taxa,sweep,sep='-')
  # is sweep dominated by a country, and also check sweep direction (enrich for UC or no)
  observed_freq <- table(sweep_list$country)
  if (length(observed_freq)>1){
    new_issweep_geo_p_unmerged_3b_UC[[s]]=c(chisq.test(observed_freq)$p.value,new_issweep_p_df_unmerged_3b_UC$coef[s]>0)  #here needs to be >0 because the correlation is for UC
  }else{
    new_issweep_geo_p_unmerged_3b_UC[[s]]=c(0,new_issweep_p_df_unmerged_3b_UC$coef[s]>0) #here needs to be >0 because the correlation is for UC
  }
  names(new_issweep_geo_p_unmerged_3b_UC)[s]=paste(taxa,sweep,sep='-')
}

new_issweep_geo_p_unmerged_3b_UC_df=as.data.frame(t(bind_rows(new_issweep_geo_p_unmerged_3b_UC,.id='id')))
names(new_issweep_geo_p_unmerged_3b_UC_df)=c('geo_enriched','UC_enriched')
new_issweep_geo_p_unmerged_3b_UC_df=new_issweep_geo_p_unmerged_3b_UC_df %>%
  mutate(geo_enriched=case_when(geo_enriched<=0.05 ~ TRUE, TRUE ~ FALSE))
new_issweep_geo_p_unmerged_3b_UC_df=new_issweep_geo_p_unmerged_3b_UC_df %>%
  mutate(UC_enriched=case_when(UC_enriched=='1' ~ 'enriched', TRUE ~ 'depleted'))

new_issweep_geo_p_unmerged_3b_UC_df_nogeo=new_issweep_geo_p_unmerged_3b_UC_df %>%
  filter(geo_enriched==FALSE) %>% select(UC_enriched)
names(new_issweep_geo_p_unmerged_3b_UC_df_nogeo)='UC_sweep_NONGEO'


```

```{r}
colors=c('#4393C3','#D6604D')
p1_UC=Heatmap(new_issweep_geo_p_unmerged_3b_UC_df_nogeo,col=colors,row_names_max_width = unit(12, "cm"))
png('non_geo_sweeps_reg_UC_vs_non_clean_0.05.png',width=2000,height=5000,res=300)
print(p1_UC)
dev.off()
```
#merge IBD, UC, CD for contrast between each other

```{r}

new_issweep_geo_p_unmerged_IBD_nogeo_comb <- list(
                  new_issweep_geo_p_unmerged_3_IBD_df_nogeo, 
                  new_issweep_geo_p_unmerged_3a_CD_df_nogeo, 
                  new_issweep_geo_p_unmerged_3b_UC_df_nogeo
                   ) %>%
  lapply(rownames_to_column, var = "rowname") %>%
  reduce(full_join, by = "rowname") %>%
  column_to_rownames(var = "rowname")

new_issweep_geo_p_unmerged_IBD_nogeo_comb$SGB=sapply(strsplit(rownames(new_issweep_geo_p_unmerged_IBD_nogeo_comb),split = '-'),'[[',1)
new_issweep_geo_p_unmerged_IBD_nogeo_comb=new_issweep_geo_p_unmerged_IBD_nogeo_comb[order(new_issweep_geo_p_unmerged_IBD_nogeo_comb$SGB), ]

new_issweep_geo_p_unmerged_IBD_nogeo_comb_IBDs=new_issweep_geo_p_unmerged_IBD_nogeo_comb %>% filter(!is.na(IBD_sweep_NONGEO)) %>% filter(!is.na(CD_sweep_NONGEO)|!is.na(UC_sweep_NONGEO))%>% arrange(CD_sweep_NONGEO)

names(new_issweep_geo_p_unmerged_IBD_nogeo_comb_IBDs)=c('IBD','CD','UC','SGB')
new_issweep_geo_p_unmerged_IBD_nogeo_comb_IBDs_1=new_issweep_geo_p_unmerged_IBD_nogeo_comb_IBDs[,c(1:3)]

SGB_list=unique(new_issweep_geo_p_unmerged_IBD_nogeo_comb_IBDs$SGB)
SGB_list=c("s__Coprococcus_comes_SGB4577","s__Bacteroides_intestinalis_SGB1846",
           "s__Hungatella_hathewayi_SGB4742","s__Bacteroides_fragilis_SGB1855",
           "s__Bacteroides_ovatus_SGB1871","s__Bacteroides_xylanisolvens_SGB1867",
           "s__Bacteroides_stercoris_SGB1830",
           "s__Bacteroides_uniformis_SGB1836",
           "s__Bacteroides_eggerthii_SGB1829",
           "s__Phocaeicola_vulgatus_SGB1814" ,"s__Lacrimispora_celerecrescens_SGB4868")

colors2=c('#4393C3','#D6604D','white')
#p2_IBDp=Heatmap(new_issweep_geo_p_unmerged_IBD_nogeo_comb_IBDs_1,col=colors2,na_col = "white",
#           rect_gp = gpar(col = "grey60", lwd = 1),
#           row_split = factor(new_issweep_geo_p_unmerged_IBD_nogeo_comb_IBDs$SGB,levels=SGB_list),
#           show_row_names = F,
#           row_title_rot = 0,
#           column_names_gp = grid::gpar(fontsize = 18),
#           column_title_rot = 00)

#png('non_geo_sweeps_IBDp_vs_non_compare_clean_0.05.png',width=2500,height=2500,res=300)
#print(p2_IBDp)
#dev.off()

new_issweep_geo_p_unmerged_IBD_nogeo_comb_single=new_issweep_geo_p_unmerged_IBD_nogeo_comb %>% filter(rowSums(is.na(.)) == 2)%>% arrange(IBD_sweep_NONGEO,CD_sweep_NONGEO)
names(new_issweep_geo_p_unmerged_IBD_nogeo_comb_single)=c('IBD','CD','UC','SGB')


SGB_list=unique(new_issweep_geo_p_unmerged_IBD_nogeo_comb_single$SGB)


new_issweep_geo_p_unmerged_IBD_nogeo_comb_single_1=new_issweep_geo_p_unmerged_IBD_nogeo_comb_single[,c(1:3)]

colors2=c('#4393C3','#D6604D','white')
p2_IBDs=Heatmap(new_issweep_geo_p_unmerged_IBD_nogeo_comb_single_1,col=colors2,na_col = "white",
           rect_gp = gpar(col = "grey60", lwd = 1),
           row_split = factor(new_issweep_geo_p_unmerged_IBD_nogeo_comb_single$SGB,levels=SGB_list),
           show_row_names = F,
           row_title_rot = 0,
           column_names_gp = grid::gpar(fontsize = 18))

png('non_geo_sweeps_IBDs_vs_non_compare_clean_0.05.png',width=2000,height=2500,res=300)
print(p2_IBDs)
dev.off()

```
## T2D metadata sweep merge 
```{r}
#this is for T2D
newtotalsweep_merged_list_4=lapply(newtotalsweep_merged_list,function(x)
  merge(baseline_data_T2D,x,by.x='sample_id',by.y='V1') %>% select(sample_id,study_name,subject_id,id,disease,disease_subtype,age,age_category,country,family, non_westernized,antibiotics_current_use))
  
newtotalsweep_merged_list_4=lapply(newtotalsweep_merged_list_4,function(x)
  x %>% 
#    mutate(family= ifelse(is.na(family),subject_id, family)) %>%
#    filter(!(age_category %in% c('newborn','child','schoolage'))) %>%
    group_by(id,family) %>% slice(1) %>% ungroup() %>% group_by(id) %>% 
    filter(n()>2) %>% ungroup())

newtotalsweep_merged_list_4=Filter(function(x) nrow(x) > 0, newtotalsweep_merged_list_4)

# what are all the metagenomes that have this SGB and can be found in the disease subset we are looking at?
newtotalstrain_list_4=newtotalstrain_list[names(newtotalsweep_merged_list_4)]

newtotalstrain_list_4=lapply(newtotalstrain_list_4,function(x)
  merge(baseline_data_T2D,x,by.x='sample_id',by.y='V1') %>% select(sample_id,subject_id,disease,disease_subtype,age,age_category,country,family, non_westernized,antibiotics_current_use)) 

newtotalstrain_list_4=mapply(function (x,y) x %>%
  group_by(family) %>%
  group_modify(~ {
    if (any(.x$sample_id %in% y$sample_id)) {
      .x %>% filter(sample_id %in% y$sample_id) %>% slice_sample(n = 1)
    } else {
      .x %>% slice_sample(n = 1)
    }
  }) %>%
  ungroup(),
  newtotalstrain_list_4,
  newtotalsweep_merged_list_4,
  SIMPLIFY = F
)

```

## T2D associations

```{r}

newtotalsweep_unmerged_list_4=lapply(newtotalsweep_merged_list_4,
                                     function(x) split(x,f=x$id))
new_issweep_list_unmerged_4=mapply(function(x,y) 
  lapply(x,function(z) y=y %>% mutate(issweep=case_when(sample_id %in% z$sample_id ~ 'sweep', TRUE~ 'notsweep'))),
  newtotalsweep_unmerged_list_4,newtotalstrain_list_4,SIMPLIFY = F)


new_issweep_p_unmerged_4=list()
new_issweep_R2_unmerged_4=list()

#i=17 (for B.frag)

for (i in 1:length(new_issweep_list_unmerged_4)){
  new_issweep_list_unmerged_4s=lapply(new_issweep_list_unmerged_4[[i]], function(x) x%>% select(disease, issweep))
  new_issweep_list_unmerged_4s=do.call(cbind,new_issweep_list_unmerged_4s)
  issweep_names=grepl('issweep',names(new_issweep_list_unmerged_4s))
  new_issweep_list_unmerged_4s=cbind(new_issweep_list_unmerged_4s[,1],
                                     new_issweep_list_unmerged_4s[issweep_names])
  names(new_issweep_list_unmerged_4s)[1]='disease'
  new_issweep_list_unmerged_4s=new_issweep_list_unmerged_4s %>% select(-where(~ all(. == "notsweep")))
  new_issweep_list_unmerged_4s=new_issweep_list_unmerged_4s %>% select(-where(~ all(. == "sweep")))
  new_issweep_list_unmerged_4s=mutate_if(new_issweep_list_unmerged_4s, is.character, as.factor)
  if (ncol(new_issweep_list_unmerged_4s)>1){
    new_issweep_list_unmerged_4s_glm=glm(disease~.,
                                         data=new_issweep_list_unmerged_4s,family='binomial')
    
    new_issweep_list_unmerged_4s_glm_null = glm(disease~1,
                                                data=new_issweep_list_unmerged_4s,family='binomial')
    
    scope = list(lower=formula(new_issweep_list_unmerged_4s_glm_null),
                 upper=formula(new_issweep_list_unmerged_4s_glm))
    
    select.fit = SignifReg(new_issweep_list_unmerged_4s_glm_null, scope = scope, direction = "forward", trace = FALSE,adjust.method = "fdr",alpha=0.1)
    R2=1-(logLik(select.fit) / logLik(new_issweep_list_unmerged_4s_glm_null))
    new_issweep_R2_unmerged_4[[i]]=R2
    
    new_issweep_tab_unmerged_4=lapply(new_issweep_list_unmerged_4[[i]], function(x) x%>% count(issweep,disease) %>% spread(disease,n,fill = 0))
    new_issweep_tab_unmerged_4=lapply(new_issweep_tab_unmerged_4, function(x) x%>% 
                                        select(-issweep))
    
    df_psweep=as.data.frame(select.fit$coefficients)
    df_psweep$sweep=row.names(df_psweep)
    rownames(df_psweep)=NULL
    names(df_psweep)[1]='coef'
    df_psweep$p.value=select.fit$steps.info$max_pvalue
    df_psweep$sweep=gsub('.issweepsweep','',df_psweep$sweep)
    df_psweep=df_psweep %>% filter(!is.na(p.value))
    if (nrow(df_psweep)>=1){
      for (s in 1:nrow(df_psweep)){
        ss=df_psweep$sweep[s]
        df_psweep$fp.value[s]=fisher.test(new_issweep_tab_unmerged_4[ss][[1]])$p.value
      }
      df_psweep$p.adjust=p.adjust(df_psweep$fp.value,method='fdr')
    }
  }else{
    df_psweep=data.frame(coef=0,sweep=NA,p.value=NA,fp.value=NA,p.adjust=NA)
  }
  
  new_issweep_p_unmerged_4[[i]]=df_psweep
}

names(new_issweep_p_unmerged_4)=names(new_issweep_list_unmerged_4)
new_issweep_p_df_unmerged_4=bind_rows(new_issweep_p_unmerged_4,.id='SGB')
new_issweep_p_df_unmerged_4=new_issweep_p_df_unmerged_4 %>% filter(!is.na(p.value))
T2D_list=new_issweep_p_df_unmerged_4$p.value<0.05
new_issweep_p_df_unmerged_4_T2D=new_issweep_p_df_unmerged_4[T2D_list,]

#filter for commensals
new_issweep_p_df_unmerged_4_T2D=new_issweep_p_df_unmerged_4_T2D %>% 
  mutate(class=case_when(SGB %in% SGB_class$pathogen ~ 'pathogens',
                         SGB %in% SGB_class$probiotics ~ 'probiotics',
                         TRUE ~ 'commensals'))
new_issweep_p_df_unmerged_4_T2D=new_issweep_p_df_unmerged_4_T2D %>% 
  filter(class=='commensals')


new_issweep_list_unmerged_4_T2D=list()
new_issweep_geo_p_unmerged_4_T2D=list()
for (s in 1:nrow(new_issweep_p_df_unmerged_4_T2D)){
  taxa=new_issweep_p_df_unmerged_4_T2D$SGB[s]
  sweep=new_issweep_p_df_unmerged_4_T2D$sweep[s]
  
  stt=new_issweep_list_unmerged_4[taxa]
  sweep_list=stt[[1]][sweep][[1]] %>% filter(issweep=='sweep')
  new_issweep_list_unmerged_4_T2D[[s]]=sweep_list
  names(new_issweep_list_unmerged_4_T2D)[s]=paste(taxa,sweep,sep='-')
  # is sweep dominated by a country, and also check sweep direction (enrich for T2D or no)
  observed_freq <- table(sweep_list$country)
  if (length(observed_freq)>1){
    new_issweep_geo_p_unmerged_4_T2D[[s]]=c(chisq.test(observed_freq)$p.value,new_issweep_p_df_unmerged_4_T2D$coef[s]>0)  #here needs to be >0 because the correlation is for T2D
  }else{
    new_issweep_geo_p_unmerged_4_T2D[[s]]=c(0,new_issweep_p_df_unmerged_4_T2D$coef[s]>0) 
  }
  names(new_issweep_geo_p_unmerged_4_T2D)[s]=paste(taxa,sweep,sep='-')
}

new_issweep_geo_p_unmerged_4_T2D_df=as.data.frame(t(bind_rows(new_issweep_geo_p_unmerged_4_T2D,.id='id')))
names(new_issweep_geo_p_unmerged_4_T2D_df)=c('geo_enriched','T2D_enriched')
new_issweep_geo_p_unmerged_4_T2D_df=new_issweep_geo_p_unmerged_4_T2D_df %>%
  mutate(geo_enriched=case_when(geo_enriched<=0.05 ~ TRUE, TRUE ~ FALSE))
new_issweep_geo_p_unmerged_4_T2D_df=new_issweep_geo_p_unmerged_4_T2D_df %>%
  mutate(T2D_enriched=case_when(T2D_enriched=='1' ~ 'enriched', TRUE ~ 'depleted'))

new_issweep_geo_p_unmerged_4_T2D_df_nogeo=new_issweep_geo_p_unmerged_4_T2D_df %>%
  filter(geo_enriched==FALSE) %>% select(T2D_enriched)
names(new_issweep_geo_p_unmerged_4_T2D_df_nogeo)='T2D_sweep_NONGEO'

```


```{r}
colors=c('#4393C3','#D6604D')
p1_T2D=Heatmap(new_issweep_geo_p_unmerged_4_T2D_df_nogeo,col=colors,row_names_max_width = unit(12, "cm"))
png('non_geo_sweeps_reg_T2D_vs_non_clean_0.05.png',width=2000,height=5000,res=300)
print(p1_T2D)
dev.off()

```

#merge the data frames with p values
```{r}
new_issweep_p_df_unmerged_1$disease='senior'
new_issweep_p_df_unmerged_2$disease='CRC'
new_issweep_p_df_unmerged_3a$disease='CD'
new_issweep_p_df_unmerged_3b$disease='UC'
new_issweep_p_df_unmerged_4$disease='T2D'
new_issweep_p_df_unmerged_2$coef=-(new_issweep_p_df_unmerged_2$coef)
new_issweep_p_df_unmerged_3a$coef=-(new_issweep_p_df_unmerged_3a$coef)
new_issweep_p_df_unmerged_all=bind_rows(new_issweep_p_df_unmerged_1,new_issweep_p_df_unmerged_2,new_issweep_p_df_unmerged_3a,
                                        new_issweep_p_df_unmerged_3b,new_issweep_p_df_unmerged_4)
 new_issweep_p_df_unmerged_all= new_issweep_p_df_unmerged_all%>% 
  mutate(class=case_when(SGB %in% SGB_class$pathogen ~ 'pathogens',
                                SGB %in% SGB_class$probiotics ~ 'probiotics',
                                TRUE ~ 'commensals')) %>% filter(class=='commensals')

new_issweep_p_df_unmerged_all$combined=paste(new_issweep_p_df_unmerged_all$SGB,new_issweep_p_df_unmerged_all$sweep,sep = '-')
siglist=bind_rows(new_issweep_geo_p_unmerged_1_age_df_nogeo,new_issweep_geo_p_unmerged_2_CRC_df_nogeo,
                  new_issweep_geo_p_unmerged_3a_CD_df_nogeo,new_issweep_geo_p_unmerged_3b_UC_df_nogeo,new_issweep_geo_p_unmerged_4_T2D_df_nogeo)

geolist=bind_rows(new_issweep_geo_p_unmerged_1_age_df,new_issweep_geo_p_unmerged_2_CRC_df,
                  new_issweep_geo_p_unmerged_3a_CD_df,new_issweep_geo_p_unmerged_3b_UC_df,new_issweep_geo_p_unmerged_4_T2D_df) %>% filter(geo_enriched=='TRUE')

new_issweep_p_df_unmerged_all=new_issweep_p_df_unmerged_all %>% mutate(association=case_when((combined %in% rownames(siglist) & p.value<0.05 & coef > 0) ~ 'positive',                                                                                              (combined %in% rownames(geolist) & p.value<0.05 & coef > 0) ~ 'positive (regional)',
(combined %in% rownames(siglist) & p.value<0.05 & coef < 0) ~ 'negative', (combined %in% rownames(geolist) & p.value<0.05 & coef < 0) ~ 'negative (regional)', TRUE ~ 'not significant'))

#new_issweep_p_df_unmerged_all=new_issweep_p_df_unmerged_all %>% mutate(disease=case_when(disease=="advanced age (65+)" ~ "advanced age", TRUE ~ disease))

new_issweep_p_df_unmerged_all$disease=gsub("senior", 'advanced age', new_issweep_p_df_unmerged_all$disease)
pv=ggplot(data = new_issweep_p_df_unmerged_all, aes(x = coef, y = -log10(p.value),shape=disease,col=association)) +
         geom_point(size=2)+theme_bw()+scale_color_manual(values = c('#2171b5','#bdd7e7','grey80','#900000','#ffb09c'),name='Association')+
  scale_shape_manual(values = c(15,2,19,4,17),name='Host state')+
  theme(legend.position = "bottom")+
   guides(shape = guide_legend(nrow = 5, byrow = TRUE,title.position = "top"),
          color=guide_legend(nrow = 5, byrow = TRUE,title.position = "top"))+xlab('Coefficent')+
  theme(text = element_text(size = 16))

png('volcano_diseases.png',res=300,width=1200,height = 1600)
pv
dev.off()

#find sweeps that are positively associated with disease and also have high coefficient

new_issweep_p_df_unmerged_all_ps=new_issweep_p_df_unmerged_all %>% filter(association=='positive')

#plot these significant sweeps so we can visualize the enrichment
all_datasets_combined=read.csv('all_datasets_combined.csv')
newtotalsweep_merged_list_all=lapply(newtotalsweep_merged_list,function(x)
  merge(all_datasets_combined,x,by.x='sample_id',by.y='V1') %>% select(sample_id,subject_id,id,disease,disease_subtype,age,age_category,country,family,non_westernized,antibiotics_current_use))

#this step makes family either family or subject id, so then if we slice for family,
#we end up slicing for both subject and family
newtotalsweep_merged_list_all=lapply(newtotalsweep_merged_list_all,function(x)
  x %>% mutate(family= ifelse(is.na(family),subject_id, family)) %>%
#    filter(!(age_category %in% c('newborn','child','schoolage'))) %>%
#    filter(disease=='healthy') %>%    #test what happens if I only consider old age and no disease
#    filter(antibiotics_current_use=='no') %>%
    group_by(id,family) %>% slice(1) %>% ungroup() %>% group_by(id) %>% 
    filter(n()>2) %>% ungroup())

newtotalsweep_merged_list_all=Filter(function(x) nrow(x) > 0, newtotalsweep_merged_list_all)

newtotalsweep_merged_list_all_SGBs=bind_rows(newtotalsweep_merged_list_all,.id='SGB')
newtotalsweep_merged_list_all_SGBs$combined=paste(newtotalsweep_merged_list_all_SGBs$SGB,newtotalsweep_merged_list_all_SGBs$id,sep='-')

newtotalsweep_merged_list_all_SGBs_select=merge(newtotalsweep_merged_list_all_SGBs,new_issweep_p_df_unmerged_all_ps,by='combined')   #so disease.x is the actual disease the sample is associated with, while disease.y is the "disease habitat" 

newtotalsweep_merged_list_all_SGBs_n=newtotalsweep_merged_list_all_SGBs %>% group_by(SGB) %>% count(disease) 


newtotalsweep_merged_list_all_SGBs_select_n=newtotalsweep_merged_list_all_SGBs_select %>% group_by(SGB.x,id,combined,disease.y) %>% count(id, disease.x) %>% ungroup()  

sgb_lookup <- newtotalsweep_merged_list_all_SGBs_select_n %>%
  distinct(SGB.x,id,combined,disease.y)
newtotalsweep_merged_list_all_SGBs_select_n = newtotalsweep_merged_list_all_SGBs_select_n %>% complete(combined,disease.x, fill = list(n = 0)) %>%
  select(combined,n,disease.x) %>% left_join(sgb_lookup,by='combined') %>% left_join(newtotalsweep_merged_list_all_SGBs_n, by = c("SGB.x" = "SGB", "disease.x" = "disease")) %>% mutate(n_norm=n.x/n.y)   #nnorm normalizes against the distribution of disease in each sweep against the disease distribution in all sweeps in the entire SGB 

# Group and collapse

newtotalsweep_merged_list_all_SGBs_select_n=newtotalsweep_merged_list_all_SGBs_select_n %>%
  group_by(SGB.x,id) %>%
  mutate(diseases= paste(unique(disease.y), collapse = ","),
    disease_sweep = paste0(diseases,' (',id,')'))  %>%
  ungroup() 


newtotalsweep_merged_list_all_SGBs_select_n$disease_sweep=gsub('newsweep','sweep',newtotalsweep_merged_list_all_SGBs_select_n$disease_sweep)

#disease_cols <- c("grey90", brewer.pal(length(unique(newtotalsweep_merged_list_all_SGBs_select_n$disease.x)),'Paired'))
disease_cols=c('grey90',wes_palette('Rushmore1',2), wes_palette('Darjeeling2', length(unique(newtotalsweep_merged_list_all_SGBs_select_n$disease.x)),type='continuous'))
disease_cols=disease_cols[c(1,3,6,7,5,8)]

disease_match=c("non-target"=disease_cols[1],"CRC"=disease_cols[2],'T2D'=disease_cols[3],'senior'=disease_cols[4],'CD'=disease_cols[5],'UC'=disease_cols[6])

newtotalsweep_merged_list_all_SGBs_select_n <- newtotalsweep_merged_list_all_SGBs_select_n %>%
  mutate(disease_sweep_colored = disease_sweep)
disease_labels=newtotalsweep_merged_list_all_SGBs_select_n$disease_sweep_colored
for (term in names(disease_match)) {
  disease_labels <- str_replace_all(
    disease_labels,
    fixed(term),  # fixed match, not regex
    paste0("<span style='color:", disease_match[[term]], ";'>", term, "</span>")
  )
}

newtotalsweep_merged_list_all_SGBs_select_n$disease_sweep_colored=disease_labels

# Plot
newtotalsweep_merged_list_all_SGBs_select_n$disease_sweep=as.factor(newtotalsweep_merged_list_all_SGBs_select_n$disease_sweep)
newtotalsweep_merged_list_all_SGBs_select_n= newtotalsweep_merged_list_all_SGBs_select_n %>% group_by(SGB.x) %>%
  arrange(disease_sweep) %>% ungroup() 

newtotalsweep_merged_list_all_SGBs_select_n$disease.x=gsub('senior','advanced age',newtotalsweep_merged_list_all_SGBs_select_n$disease.x)
newtotalsweep_merged_list_all_SGBs_select_n$disease.x <- factor(newtotalsweep_merged_list_all_SGBs_select_n$disease.x, 
                                                                levels = c("non-target", "CRC", "T2D","advanced age","CD","UC"))
newtotalsweep_merged_list_all_SGBs_select_n_habitat=newtotalsweep_merged_list_all_SGBs_select_n %>% group_by(combined) %>% slice(1)
#newtotalsweep_merged_list_all_SGBs_select_n_habitat$disease_sweep=factor(newtotalsweep_merged_list_all_SGBs_select_n_habitat$disease_sweep,
#                                                                        levels=c("age (65+)","CRC","T2D",'CD','UC'))
#newtotalsweep_merged_list_all_SGBs_select_n$disease_sweep=factor(newtotalsweep_merged_list_all_SGBs_select_n$disease_sweep,
#                                                                        levels=c("age (65+)","CRC","T2D",'CD','UC'))

# habitat_total=ggplot(newtotalsweep_merged_list_all_SGBs_select_n, aes(x = disease.y, y = n_norm, fill = disease.x)) +
#   facet_wrap(~SGB.x,scales = "free_y", drop = TRUE)+
#   geom_bar(position = "fill",stat = "identity") +
#   scale_fill_manual(values=disease_cols)+
#   scale_y_continuous(expand = c(0,0),labels = scales::percent_format()) +
#  # scale_x_discrete(labels = newtotalsweep_merged_list_all_SGBs_select_n_habitat$disease.y)+
#   theme_minimal(base_size = 14) +
#   labs(
#     x    = "Associated host state",
#     y    = "% samples in sweep",
#     fill = "Disease"
#   )+ coord_flip()

newtotalsweep_merged_list_all_SGBs_select_n$facet_label=gsub('s__', '*', newtotalsweep_merged_list_all_SGBs_select_n$SGB.x)
newtotalsweep_merged_list_all_SGBs_select_n$facet_label=gsub('_', ' ', newtotalsweep_merged_list_all_SGBs_select_n$facet_label)
newtotalsweep_merged_list_all_SGBs_select_n$facet_label=gsub(' SGB','*<br> (SGB',newtotalsweep_merged_list_all_SGBs_select_n$facet_label)
newtotalsweep_merged_list_all_SGBs_select_n$facet_label=paste0(newtotalsweep_merged_list_all_SGBs_select_n$facet_label,')')

habitat_total_2=ggplot(newtotalsweep_merged_list_all_SGBs_select_n, aes(y = disease_sweep_colored, x = n_norm, fill = disease.x)) +
  facet_wrap(~facet_label,scales = "free_y", drop = TRUE,ncol=3)+
  geom_bar(position = "fill",stat = "identity") +
  scale_fill_manual(values=disease_cols)+
  scale_x_continuous(expand = c(0,0),labels = scales::percent_format()) +
   theme_minimal() +
  theme(axis.text.y = ggtext::element_markdown(size = 12),
        axis.text.x = element_text(size = 12),
        strip.text = ggtext::element_markdown(size = 12),
        axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16))+
  labs(
    y    = "Associated host state",
    x    = "% samples in sweep",
    fill = "Host state"
  )

# png('habitat_total.png',res=300,width = 6000,height = 3000)
# habitat_total
# dev.off()

png('habitat_total_sweep.png',res=300,width = 4500,height = 4000)
habitat_total_2
dev.off()

newtotalsweep_merged_list_all_SGBs_select_n_BU=newtotalsweep_merged_list_all_SGBs_select_n %>% filter(SGB.x=='s__Bacteroides_uniformis_SGB1836') 
newtotalsweep_merged_list_all_SGBs_select_n_BC=newtotalsweep_merged_list_all_SGBs_select_n %>% filter(SGB.x=='s__Bacteroides_caccae_SGB1877') 


newtotalsweep_merged_list_all_SGBs_select_n_BU$disease.x <- factor(newtotalsweep_merged_list_all_SGBs_select_n_BU$disease.x, 
                                                                levels = c("non-target", "CRC", "T2D","advanced age","CD","UC"))
# newtotalsweep_merged_list_all_SGBs_select_n_BU_habitat=newtotalsweep_merged_list_all_SGBs_select_n_BU %>% group_by(combined) %>% slice(1)
# newtotalsweep_merged_list_all_SGBs_select_n_BU_habitat$disease.y=factor(newtotalsweep_merged_list_all_SGBs_select_n_BU_habitat$disease.y,
#                                                                         levels=c("age (65+)","CRC","T2D",'CD','UC'))
# 
# newtotalsweep_merged_list_all_SGBs_select_n_BU$disease.y=factor(newtotalsweep_merged_list_all_SGBs_select_n_BU$disease.y,
#                                                                         levels=c("age (65+)","CRC","T2D",'CD','UC'))

# habitat_BU=ggplot(newtotalsweep_merged_list_all_SGBs_select_n_BU, aes(x = disease.y, y = n_norm, fill = disease.x)) +
#   facet_wrap(~SGB.x,scales = "free_x", drop = TRUE)+
#   geom_bar(position = "fill",stat = "identity") +
#   scale_fill_manual(values=disease_cols)+
#   scale_y_continuous(expand = c(0,0),labels = scales::percent_format()) +
#  # scale_x_discrete(labels = newtotalsweep_merged_list_all_SGBs_select_n_habitat$disease.y)+
#   theme_minimal(base_size = 14) +
#   labs(
#     x    = "Associated host state",
#     y    = "% samples in sweep",
#     fill = "Disease"
#   )+ 
#   coord_flip()
# 
# png('habitat_BU.png',res=300,width = 2000,height = 1000)
# habitat_BU
# dev.off()
newtotalsweep_merged_list_all_SGBs_select_n_BU$disease_sweep_colored=as.factor(newtotalsweep_merged_list_all_SGBs_select_n_BU$disease_sweep_colored)
#newtotalsweep_merged_list_all_SGBs_select_n_BU$disease_sweep=factor(newtotalsweep_merged_list_all_SGBs_select_n_BU$disease_sweep,levels=                  #                                                    levels(newtotalsweep_merged_list_all_SGBs_select_n_BU$disease_sweep)[c(1,4:9,2:3,10:12)])

newtotalsweep_merged_list_all_SGBs_select_n_BU$facet_label=gsub('s__', '*', newtotalsweep_merged_list_all_SGBs_select_n_BU$SGB.x)
newtotalsweep_merged_list_all_SGBs_select_n_BU$facet_label=gsub('_', ' ', newtotalsweep_merged_list_all_SGBs_select_n_BU$facet_label)
newtotalsweep_merged_list_all_SGBs_select_n_BU$facet_label=gsub(' SGB','* (SGB',newtotalsweep_merged_list_all_SGBs_select_n_BU$facet_label)
newtotalsweep_merged_list_all_SGBs_select_n_BU$facet_label=paste0(newtotalsweep_merged_list_all_SGBs_select_n_BU$facet_label,')')

habitat_BU_2=ggplot(newtotalsweep_merged_list_all_SGBs_select_n_BU, aes(y = disease_sweep_colored, x = n_norm, fill = disease.x)) +
  facet_wrap(~ facet_label,scales = "free_y", drop = TRUE)+
  geom_bar(position = "fill",stat = "identity") +
  theme_minimal(base_size = 14) +
  scale_fill_manual(values=disease_cols)+
  scale_x_continuous(expand = c(0,0),labels = scales::percent_format()) +
  theme(axis.text.y = ggtext::element_markdown(size = 14),
        strip.text = ggtext::element_markdown(size = 14))+
 # scale_x_discrete(labels = newtotalsweep_merged_list_all_SGBs_select_n_habitat$disease.y)+
  labs(
    y    = "Associated host state",
    x    = "% samples in sweep",
    fill = "Host state"
  )



png('habitat_BU_sweep.png',res=300,width = 2500,height = 1000)
habitat_BU_2
dev.off()

newtotalsweep_merged_list_all_SGBs_select_n_BC$disease.x <- factor(newtotalsweep_merged_list_all_SGBs_select_n_BC$disease.x, 
                                                                levels = c("non-target", "CRC", "T2D","advanced age","CD","UC"))
# newtotalsweep_merged_list_all_SGBs_select_n_BC_habitat=newtotalsweep_merged_list_all_SGBs_select_n_BC %>% group_by(combined) %>% slice(1)
# newtotalsweep_merged_list_all_SGBs_select_n_BC_habitat$disease.y=factor(newtotalsweep_merged_list_all_SGBs_select_n_BC_habitat$disease.y,
#                                                                         levels=c("age (65+)","CRC","T2D",'CD','UC'))
# 
# newtotalsweep_merged_list_all_SGBs_select_n_BC$disease.y=factor(newtotalsweep_merged_list_all_SGBs_select_n_BC$disease.y,
#                                                                         levels=c("age (65+)","CRC","T2D",'CD','UC'))
# newtotalsweep_merged_list_all_SGBs_select_n_BC$disease_sweep=as.factor(newtotalsweep_merged_list_all_SGBs_select_n_BC$disease_sweep)
# newtotalsweep_merged_list_all_SGBs_select_n_BC$disease_sweep=factor(newtotalsweep_merged_list_all_SGBs_select_n_BC$disease_sweep,levels = levels(newtotalsweep_merged_list_all_SGBs_select_n_BC$disease_sweep)[c(1,3:4,2,5:7)])

# habitat_BC=ggplot(newtotalsweep_merged_list_all_SGBs_select_n_BC, aes(x = disease.y, y = n_norm, fill = disease.x)) +
#   facet_wrap(~SGB.x,scales = "free_x", drop = TRUE)+
#   geom_bar(position = "fill",stat = "identity") +
#   scale_fill_manual(values=disease_cols)+
#   scale_y_continuous(expand = c(0,0),labels = scales::percent_format()) +
#  # scale_x_discrete(labels = newtotalsweep_merged_list_all_SGBs_select_n_habitat$disease.y)+
#   theme_minimal(base_size = 14) +
#   labs(
#     x    = "Associated host state",
#     y    = "% samples in sweep",
#     fill = "Disease"
#   )+ coord_flip()
# 
# png('habitat_BC.png',res=300,width = 2000,height = 1000)
# habitat_BC
# dev.off()

habitat_BC_2=ggplot(newtotalsweep_merged_list_all_SGBs_select_n_BC, aes(x = disease_sweep, y = n_norm, fill = disease.x)) +
  facet_wrap(~SGB.x,scales = "free_x", drop = TRUE)+
  geom_bar(position = "fill",stat = "identity") +
  scale_fill_manual(values=disease_cols)+
  scale_y_continuous(expand = c(0,0),labels = scales::percent_format()) +
 # scale_x_discrete(labels = newtotalsweep_merged_list_all_SGBs_select_n_habitat$disease.y)+
  theme_minimal(base_size = 14) +
  labs(
    x    = "Associated host state",
    y    = "% samples in sweep",
    fill = "Disease"
  )+ coord_flip()

png('habitat_BC_sweep.png',res=300,width = 2000,height = 1000)
habitat_BC_2
dev.off()


```

#age,CRC,IBD,T2D merge

```{r}
new_issweep_geo_p_unmerged_nogeo_comb <- list(new_issweep_geo_p_unmerged_1_age_df_nogeo, 
                                              new_issweep_geo_p_unmerged_4_T2D_df_nogeo,
                  new_issweep_geo_p_unmerged_2_CRC_df_nogeo, 
                  new_issweep_geo_p_unmerged_3_IBD_df_nogeo
                   ) %>%
  lapply(rownames_to_column, var = "rowname") %>%
  reduce(full_join, by = "rowname") %>%
  column_to_rownames(var = "rowname")

new_issweep_geo_p_unmerged_nogeo_comb$SGB=sapply(strsplit(rownames(new_issweep_geo_p_unmerged_nogeo_comb),split = '-'),'[[',1)

new_issweep_geo_p_unmerged_nogeo_comb_dircount=new_issweep_geo_p_unmerged_nogeo_comb %>% group_by(SGB) %>% 
  summarize(unique_age = n_distinct(age_sweep_NONGEO,na.rm=T),unique_CRC=n_distinct(CRC_sweep_NONGEO,na.rm=T),
            unique_IBD = n_distinct(IBD_sweep_NONGEO,na.rm=T),unique_T2D=n_distinct(IBD_sweep_NONGEO,na.rm=T))

          
oppo_SGB=new_issweep_geo_p_unmerged_nogeo_comb_dircount %>% filter(unique_age>1|unique_CRC>1|unique_IBD>1|unique_T2D>1)
new_issweep_geo_p_unmerged_nogeo_comb_oppo=new_issweep_geo_p_unmerged_nogeo_comb %>% filter(SGB %in% oppo_SGB$SGB) 

names(new_issweep_geo_p_unmerged_nogeo_comb_oppo)=c('age','T2D','CRC','IBD','SGB')
new_issweep_geo_p_unmerged_nogeo_comb_oppo_1=new_issweep_geo_p_unmerged_nogeo_comb_oppo[,c(1:4)]

write.csv(new_issweep_geo_p_unmerged_nogeo_comb, 'new_issweep_geo_p_unmerged_nogeo_comb_clean_0.05.csv')
write.csv(new_issweep_geo_p_unmerged_nogeo_comb_oppo, 'new_issweep_geo_p_unmerged_nogeo_comb_oppo_clean_0.05.csv')

colors2=c('#4393C3','#D6604D','white')
col_mapping <- c("enriched" = '#D6604D', "depleted" = "#4393C3")
p2=Heatmap(new_issweep_geo_p_unmerged_nogeo_comb_oppo_1,col=col_mapping,na_col = "white",
           column_gap = unit(5, "mm"),
           column_split=c('age','T2D','CRC','IBD'),
           rect_gp = gpar(col = "grey60", lwd = 1),
           row_split = factor(new_issweep_geo_p_unmerged_nogeo_comb_oppo$SGB),
           show_row_names = F,
           row_title_rot = 0)

png('non_geo_sweeps_reg_all_bidirectional_clean_0.05.png',width=2000,height=1500,res=300)
print(p2)
dev.off()

new_issweep_geo_p_unmerged_nogeo_comb_2cond=new_issweep_geo_p_unmerged_nogeo_comb %>% filter(rowSums(is.na(.)) <= 2)
names(new_issweep_geo_p_unmerged_nogeo_comb_2cond)=c('age','T2D','CRC','IBD','SGB')
new_issweep_geo_p_unmerged_nogeo_comb_2cond_1=new_issweep_geo_p_unmerged_nogeo_comb_2cond[,c(1:4)]

colors2=c('#4393C3','#D6604D','white')
col_mapping <- c("enriched" = '#D6604D', "depleted" = "#4393C3")
p2=Heatmap(new_issweep_geo_p_unmerged_nogeo_comb_2cond_1,col=col_mapping,na_col = "white",
           column_gap = unit(5, "mm"),
           column_split=c('age','T2D','CRC','IBD'),
           rect_gp = gpar(col = "grey60", lwd = 1),
           row_split = factor(new_issweep_geo_p_unmerged_nogeo_comb_2cond$SGB),
           show_row_names = F,
           row_title_rot = 0)

png('non_geo_sweeps_reg_all_2cond_clean_0.05.png',width=2000,height=1500,res=300)
print(p2)
dev.off()


```


# associate sweep - isolate match
```{r}

#check how much associated sweeps may overlap with isolate based sweeps
new_issweep_geo_p_unmerged_nogeo_comb6 <- list(new_issweep_geo_p_unmerged_1_age_df_nogeo, 
                                              new_issweep_geo_p_unmerged_4_T2D_df_nogeo,
                  new_issweep_geo_p_unmerged_2_CRC_df_nogeo, 
   
                  new_issweep_geo_p_unmerged_3a_CD_df_nogeo,
                  new_issweep_geo_p_unmerged_3b_UC_df_nogeo
                   ) %>%
  lapply(rownames_to_column, var = "rowname") %>%
  reduce(full_join, by = "rowname") %>%
  column_to_rownames(var = "rowname")


associated_sweeps=row.names(new_issweep_geo_p_unmerged_nogeo_comb6)
associated_sweeps_strains=list()
for (s in 1:length(associated_sweeps)){
  st=strsplit(associated_sweeps[s],split='-')[[1]][1]
  ss=strsplit(associated_sweeps[s],split='-')[[1]][2]
  associated_sweeps_strains[[s]]=as.data.frame(newtotalsweep_list[st][[1]][ss])
}
names(associated_sweeps_strains)=associated_sweeps

associated_sweeps_strains_all=bind_rows(associated_sweeps_strains,.id = 'newsweep')

#read in isolate based sweeps
isolate_sweeps_table=Sys.glob('../../instrain/confirmed_select_lists/*.txt')

isolate_sweeps_list=lapply(isolate_sweeps_table, function(x) read.table(x))
names(isolate_sweeps_list)=unlist(lapply(isolate_sweeps_table,function(x) gsub('../../instrain/confirmed_select_lists/','',x)))
names(isolate_sweeps_list)=gsub('_consensus_sweep.txt','',names(isolate_sweeps_list))
isolate_sweeps_all=bind_rows(isolate_sweeps_list,.id='sweep')
isolate_sweeps_isolate_only=isolate_sweeps_all %>% filter(V1 %in% metadata_isolate$Strain)
associated_sweeps_strains_overlap=merge(associated_sweeps_strains_all,isolate_sweeps_isolate_only,by='V1') %>% arrange(newsweep)
associated_sweeps_strains_overlap=merge(associated_sweeps_strains_overlap,new_issweep_geo_p_unmerged_nogeo_comb6,by.x='newsweep',by.y=0)

associated_sweeps_strains_overlap_list=split(associated_sweeps_strains_overlap,f=associated_sweeps_strains_overlap$newsweep)
for (i in 1:length(associated_sweeps_strains_overlap_list)){
  s=unique(associated_sweeps_strains_overlap_list[[i]]$sweep)
  ni=sum(isolate_sweeps_isolate_only$sweep==s)
  r=nrow(associated_sweeps_strains_overlap_list[[i]])/ni
  associated_sweeps_strains_overlap_list[[i]]$newsweepinsweep=r
}

associated_sweeps_strains_overlap=bind_rows(associated_sweeps_strains_overlap_list)

write.csv(associated_sweeps_strains_overlap,'associated_sweeps_strains_overlap.csv',row.names = F)

select_sweeps=associated_sweeps_strains_overlap %>% filter(newsweepinsweep>0.5) %>% select(sweep) %>% unique()
select_sweeps$sweep=gsub('diffH','',select_sweeps$sweep)
select_sweeps$enrichment=c('UC_enriched_but_seperated','CRC enriched','UC enriched','CD depleted','UC_depleted','CD depleted',
                           
                           'CD_enriched','CD_UC_enriched','UC_enriched','CD_enriched')

unique_genes=read.csv('../../functional_analysis_by_SGB/sweep_differentiating_genes_clean.csv')
positive_genes=read.csv('../../functional_analysis_by_SGB/genes_under_positive_selection.csv')
positive_genes$sweep=sapply(strsplit(positive_genes$query,split='_consensus'),'[[',1)
positive_genes$sweep=gsub('diffH','',positive_genes$sweep)
unique_genes_select=merge(unique_genes,select_sweeps,by.x='Sweep',by.y='sweep')
positive_genes_select=merge(positive_genes,select_sweeps,by='sweep')
write.csv(unique_genes_select,'associated_sweeps_unique_genes.csv',row.names = F)
write.csv(positive_genes_select,'associated_sweeps_positive_genes.csv',row.names = F)

```


##merge with SGB correlations

```{r}
#read in SGB correlations and sweep correlations
age_CRC_IBD_T2D_merged_df=read.csv('age_CRC_IBD_T2D_merged_df.csv',row.names = 1)
new_issweep_geo_p_unmerged_nogeo_comb=read.csv('new_issweep_geo_p_unmerged_nogeo_comb.csv',row.names = 1)

new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_1=merge(new_issweep_geo_p_unmerged_nogeo_comb,age_CRC_IBD_T2D_merged_df,by.x='SGB',by.y=0, all.x=T)
new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_1$SGB=gsub('s__','',new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_1$SGB)


```
##age
```{r}
new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age=new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_1 %>%
  select("SGB","age","Seniority")
new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age$Seniority=as.factor(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age$Seniority)
new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age_m=new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age %>% select("age")
  
new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age$SGB=gsub('_',' ',new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age$SGB)
new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age$SGB1=sapply(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age$SGB, function(x) {
  parts <- strsplit(x, "SGB")[[1]]
  parts[[2]]=paste0("(SGB",parts[[2]],")")
  paste0(parts[[1]],parts[[2]])
})
SGB_list=unique(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age$SGB1)
SGB_list=SGB_list[c(1:3,5,4)]

#new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age_m=apply(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_m,2,as.factor)
#new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age_m=as.matrix(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_m)

ha=rowAnnotation(SGB=new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age$Seniority,
                 col=list(SGB=c("enriched" = "#D6604D", 
                                "depleted" = "#4393C3")),
                 annotation_name_gp = gpar(fontsize = 16),
                 na_col = "grey90")

p2a=Heatmap(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age_m,col=colors2,na_col = "grey90",
           rect_gp = gpar(col = "grey60", lwd = 1),
           column_split = c('age (65+)'),
           show_row_names = F,
           show_column_names = FALSE,
           cluster_column_slices = FALSE,
           cluster_row_slices = FALSE,
           row_split = factor(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_age$SGB1,
                              levels=SGB_list),
           row_gap = unit(10, "mm"),
           row_names_max_width = unit(12, "cm"),name='Direction',
           column_names_gp = grid::gpar(fontsize = 18),
           column_title_gp = grid::gpar(fontsize = 20),
           row_title_rot = 0,
           row_names_gp = grid::gpar(fontsize = 18),
           row_title_gp = grid::gpar(fontsize = 20),
           left_annotation = ha,
           heatmap_legend_param = list(labels_gp = grid::gpar(fontsize = 18),
                                       legend_direction = "horizontal", 
                                       nrow=1,
                                       legend_width = unit(0.9, "npc"),
                                       title_gp = grid::gpar(fontsize = 20)))

png('all_sweeps_reg_age_bidirection.png',width=2800,height=2000,res=300)
draw(p2a)
dev.off()
```

##CRC

```{r}

new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_CRC=new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_1 %>%
  select("SGB","CRC.x","CRC.y")
new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_CRC$Seniority=as.factor(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_CRC$CRC.y)
new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_CRC_m=new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_CRC %>% select("CRC.x")
  
new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_CRC$SGB=gsub('_',' ',new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_CRC$SGB)
new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_CRC$SGB1=sapply(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_CRC$SGB, function(x) {
  parts <- strsplit(x, "SGB")[[1]]
  parts[[2]]=paste0("(SGB",parts[[2]],")")
  paste0(parts[[1]],parts[[2]])
})
SGB_list=unique(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_CRC$SGB1)
SGB_list=SGB_list[c(1:3,5,4)]


ha=rowAnnotation(SGB=new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_CRC$CRC.y,
                 col=list(SGB=c("enriched" = "#D6604D", 
                                "depleted" = "#4393C3")),
                 annotation_name_gp = gpar(fontsize = 16),
                 na_col = "grey90")

p2a=Heatmap(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_CRC_m,col=colors2,na_col = "grey90",
           rect_gp = gpar(col = "grey60", lwd = 1),
           column_split = c('CRC'),
           show_row_names = F,
           show_column_names = FALSE,
           cluster_column_slices = FALSE,
           cluster_row_slices = FALSE,
           row_split = factor(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_CRC$SGB1,
                              levels=SGB_list),
           row_gap = unit(10, "mm"),
           row_names_max_width = unit(12, "cm"),name='Direction',
           column_names_gp = grid::gpar(fontsize = 18),
           column_title_gp = grid::gpar(fontsize = 20),
           row_title_rot = 0,
           row_names_gp = grid::gpar(fontsize = 18),
           row_title_gp = grid::gpar(fontsize = 20),
           left_annotation = ha,
           heatmap_legend_param = list(labels_gp = grid::gpar(fontsize = 18),
                                       legend_direction = "horizontal", 
                                       nrow=1,
                                       legend_width = unit(0.9, "npc"),
                                       title_gp = grid::gpar(fontsize = 20)))

png('all_sweeps_reg_CRC_bidirection.png',width=2800,height=2000,res=300)
draw(p2a)
dev.off()

```
##T2D

```{r}

new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_T2D=new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_1 %>%
  select("SGB","T2D.x","T2D.y")
new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_T2D$Seniority=as.factor(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_T2D$T2D.y)
new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_T2D_m=new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_T2D %>% select("T2D.x")
  
new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_T2D$SGB=gsub('_',' ',new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_T2D$SGB)
new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_T2D$SGB1=sapply(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_T2D$SGB, function(x) {
  parts <- strsplit(x, "SGB")[[1]]
  parts[[2]]=paste0("(SGB",parts[[2]],")")
  paste0(parts[[1]],parts[[2]])
})
SGB_list=unique(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_T2D$SGB1)
SGB_list=SGB_list[c(1:3,5,4)]


ha=rowAnnotation(SGB=new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_T2D$T2D.y,
                 col=list(SGB=c("enriched" = "#D6604D","depleted" = "#4393C3"
                                )),
                 annotation_name_gp = gpar(fontsize = 16),
                 na_col = "grey90")

p2a=Heatmap(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_T2D_m,col=colors2[2:3],na_col = "grey90",
           rect_gp = gpar(col = "grey60", lwd = 1),
           column_split = c('T2D'),
           show_row_names = F,
           show_column_names = FALSE,
           cluster_column_slices = FALSE,
           cluster_row_slices = FALSE,
           row_split = factor(new_issweep_geo_p_unmerged_both_df_geo_new_comb_filter_T2D$SGB1,
                              levels=SGB_list),
           row_gap = unit(10, "mm"),
           row_names_max_width = unit(12, "cm"),name='Direction',
           column_names_gp = grid::gpar(fontsize = 18),
           column_title_gp = grid::gpar(fontsize = 20),
           row_title_rot = 0,
           row_names_gp = grid::gpar(fontsize = 18),
           row_title_gp = grid::gpar(fontsize = 20),
           left_annotation = ha,
           heatmap_legend_param = list(labels_gp = grid::gpar(fontsize = 18),
                                       legend_direction = "horizontal", 
                                       nrow=1,
                                       legend_width = unit(0.9, "npc"),
                                       title_gp = grid::gpar(fontsize = 20)))

png('all_sweeps_reg_T2D_bidirection.png',width=2450,height=2000,res=300)
draw(p2a)
dev.off()

```


##general enrichment in sweeps
```{r}
# first ask if any symptoms are enriched in sweeps for age
new_issweep_list_1=mapply(function(x,y) y=y %>% mutate(issweep=case_when(sample_id %in% x$sample_id ~ 'sweep', TRUE~ 'notsweep')),
                          newtotalsweep_merged_list_1,newtotalstrain_list_1,SIMPLIFY = F)


new_issweep_p_1=list()
new_issweep_tab_list_1=list()
for (i in 1:length(new_issweep_list_1)){
  new_issweep_tab_1=new_issweep_list_1[[i]] %>% count(issweep,age_category) %>% spread(age_category,n,fill = 0)
  new_issweep_tab_1= new_issweep_tab_1 %>% 
    select(-issweep)
  new_issweep_tab_list_1[[i]]=new_issweep_tab_1
  if (nrow(new_issweep_tab_1)>1){
    new_issweep_p_1[[i]]=chisq.test(new_issweep_tab_1)$p.value
  } else {
    new_issweep_p_1[[i]]=1
  }
}

names(new_issweep_p_1)=names(new_issweep_list_1)
names(new_issweep_tab_list_1)=names(new_issweep_list_1)
new_issweep_p_df_1=t(bind_rows(new_issweep_p_1,.id='SGB'))
new_issweep_p_df_1=as.data.frame(new_issweep_p_df_1)
new_issweep_p_df_1$p.adj=p.adjust(new_issweep_p_df_1$V1,method = 'fdr')
age_list=rownames(new_issweep_p_df_1)[new_issweep_p_df_1$p.adj<0.05]
new_issweep_p_df_1[age_list,]
```


```{r}
# ask if any symptoms are enriched in sweeps, CRC
new_issweep_list_2=mapply(function(x,y) y=y %>% mutate(issweep=case_when(sample_id %in% x$sample_id ~ 'sweep', TRUE~ 'notsweep')),
                          newtotalsweep_merged_list_2,newtotalstrain_list_2,SIMPLIFY = F)

new_issweep_p_2=list()
new_issweep_tab_list_2=list()
for (i in 1:length(new_issweep_list_2)){
  new_issweep_tab_2=new_issweep_list_2[[i]] %>% count(issweep,disease) %>% spread(disease,n,fill = 0)
  new_issweep_tab_2= new_issweep_tab_2 %>% 
    select(-issweep)
  new_issweep_tab_list_2[[i]]=new_issweep_tab_2
  if (nrow(new_issweep_tab_2)>1){
    new_issweep_p_2[[i]]=chisq.test(new_issweep_tab_2)$p.value
  } else {
    new_issweep_p_2[[i]]=1
  }
}

names(new_issweep_p_2)=names(new_issweep_list_2)
names(new_issweep_tab_list_2)=names(new_issweep_list_2)
new_issweep_p_df_2=t(bind_rows(new_issweep_p_2,.id='SGB'))
new_issweep_p_df_2=as.data.frame(new_issweep_p_df_2)
new_issweep_p_df_2$p.adjust=p.adjust(new_issweep_p_df_2$V1,method='fdr')
CRC_list=rownames(new_issweep_p_df_2)[new_issweep_p_df_2$p.adjust<0.05]
new_issweep_p_df_2[CRC_list,]

```
```{r}
# ask if any symptoms are enriched in sweeps, IBD
new_issweep_list_3=mapply(function(x,y) y=y %>% mutate(issweep=case_when(sample_id %in% x$sample_id ~ 'sweep', TRUE~ 'notsweep')),
                          newtotalsweep_merged_list_3,newtotalstrain_list_3,SIMPLIFY = F)

new_issweep_p_3=list()
new_issweep_tab_list_3=list()
for (i in 1:length(new_issweep_list_3)){
  new_issweep_tab_3=new_issweep_list_3[[i]] %>% count(issweep,disease) %>% spread(disease,n,fill = 0)
  new_issweep_tab_3= new_issweep_tab_3 %>% 
    select(-issweep)
  new_issweep_tab_list_3[[i]]=new_issweep_tab_3
  if (nrow(new_issweep_tab_3)>1){
    new_issweep_p_3[[i]]=chisq.test(new_issweep_tab_3)$p.value
  } else {
    new_issweep_p_3[[i]]=1
  }
}

names(new_issweep_p_3)=names(new_issweep_list_3)
names(new_issweep_tab_list_3)=names(new_issweep_list_3)
new_issweep_p_df_3=t(bind_rows(new_issweep_p_3,.id='SGB'))
new_issweep_p_df_3=as.data.frame(new_issweep_p_df_3)
new_issweep_p_df_3$p.adjust=p.adjust(new_issweep_p_df_3$V1,method='fdr')
IBD_list=rownames(new_issweep_p_df_3)[new_issweep_p_df_3$p.adjust<0.05]
new_issweep_p_df_3[IBD_list,]

```
```

```{r}
# ask if any symptoms are enriched in sweeps, T2D
new_issweep_list_4=mapply(function(x,y) y=y %>% mutate(issweep=case_when(sample_id %in% x$sample_id ~ 'sweep', TRUE~ 'notsweep')),
                          newtotalsweep_merged_list_4,newtotalstrain_list_4,SIMPLIFY = F)


new_issweep_p_4=list()
new_issweep_tab_list_4=list()
for (i in 1:length(new_issweep_list_4)){
  new_issweep_tab_4=new_issweep_list_4[[i]] %>% count(issweep,disease) %>% spread(disease,n,fill = 0)
  new_issweep_tab_4= new_issweep_tab_4 %>% 
    select(-issweep)
  new_issweep_tab_list_4[[i]]=new_issweep_tab_4
  if (nrow(new_issweep_tab_4)>1){
    new_issweep_p_4[[i]]=chisq.test(new_issweep_tab_4)$p.value
  } else {
    new_issweep_p_4[[i]]=1
  }
}

names(new_issweep_p_4)=names(new_issweep_list_4)
names(new_issweep_tab_list_4)=names(new_issweep_list_4)
new_issweep_p_df_4=t(bind_rows(new_issweep_p_4,.id='SGB'))
new_issweep_p_df_4=as.data.frame(new_issweep_p_df_4)
new_issweep_p_df_4$p.adjust=p.adjust(new_issweep_p_df_4$V1,method='fdr')
T2D_list=rownames(new_issweep_p_df_4)[new_issweep_p_df_4$p.adjust<0.05]
new_issweep_p_df_4[T2D_list,]

```


